---
title: "Chewing Gum"
date: "`r format(Sys.time(), '%d %B, %Y (%H:%M)')`"
output:
  html_document:
    font: Source Sans Pro
    theme: cosmo
    anchor_sections: TRUE
    toc_float: 
      collapsed: true
      smooth_scroll: TRUE
    toc_depth: 3
    linkcolor: blue
  latex_engine: xelatex
  pdf_document:
    toc: yes
    toc_depth: '3'
bibliography: "bib/gum.bib"
csl: jama.csl
link-citations: yes
workflowr:
  suppress_report: false
nocite: '@*'
---

#### 

```{=html}
<style type="text/css">
  body{
  font-size: 9.5pt;
  font-family: Source Sans Pro;
}

.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
  color: #090909;
  background-color: #0909093d;
}

.lightable-classic.lightable-hover tbody tr:hover {
    background-color: #c5d4e4;
}

.btn-xs, .btn-group-xs>.btn { 
display: none; 
}

.btn-default {
display: none !important; 
}

.shortbr {
line-height: 30px;
}

<!-- .manchor { -->
<!--     display: block; -->
<!--     position: relative; -->
<!--     top: -250px; -->
<!--     visibility: hidden; -->
<!-- } -->

<!-- .manchor { -->
<!--   padding-top: 400px; -->
<!--   margin-top: -40px; -->
<!--   display: inline-block; /* required for webkit browsers */ -->
<!-- } -->

:target {  /* fix target location so caption appears */
    display: block;    
    position: relative;     
    top: 10px;
    visibility: hidden;
}

</style>

```

```{r setup, include = FALSE}
## (updated 2021/01/06 11:34) preliminaries -------------------------------
# functions
now <- function() {str_replace_all(format(Sys.time(), "%m %d %y_%H-%M"), " ", "_")}
# replace_mg <- function(a, b) {if_else(is.na(a) & !is.na(b), b, a)}

tab_inc <- function() {
  table_n <- table_n + 1
  table_n}

fig_inc <- function() {
  figure_n <- figure_n + 1
  figure_n}

table_n <- 1
figure_n <- 1
```

```{r filesRecent, include = FALSE}
## get current files; select protein refids; directory must have unique ####
source("code/readFiles_120220_wfr.R") 

# add study gum indicator "tx" -- gum_sugared, gum_sugarfree, sugared_sugarfree
study_char.dat <- left_join(study_char.dat, gum_refids, by = "refid") 
study_arm.dat  <- left_join(study_arm.dat, gum_refids, by = "refid") 

# study char gum refs 
study_char_gum.dat <- study_char.dat %>% 
  filter(refid %in% gum_refids$refid)

# study arm gum refs 
study_arm_gum.dat <- study_arm.dat %>% 
  filter(refid %in% gum_refids$refid) %>% 
  mutate(arm_tx = ifelse(refid == 2009, "unknown", arm_tx))

# change gum_refids 2009 to unknown
gum_refids <- gum_refids %>% 
  mutate(tx = ifelse(refid == 2009, "unknown", tx))

```

# **Included Studies**

<font size = 4> Table `r table_n`. Included studies according to age, surgery, and design (see [References](#references) for citations). </font>

```{r studies, echo = FALSE, include = TRUE}
## (updated 2020/12/17 20:31) studies -------------------------------------
# list of studies, age, surgical/non surgical, design
study_char_gum.dat %>% 
  mutate(
    surg = case_when(
      str_detect(surg_type, "[Gg]yn") ~ "Gyn",
      str_detect(surg_type, "Elective") ~ "Elective",
      str_detect(surg_type, "esophagog") ~ "Endoscopy",
      str_detect(surg_type, "total hip or total knee replacement") ~ "TKA/THA",
      str_detect(surg_type, "cholecystectomy") ~ "Cholecystectomy",
      str_detect(surg_type, "[Gg]ast|Bowel") ~ "Other GI",
      str_detect(surg_type, "not appli") ~ "None (healthy)",
      str_detect(surg_type, "oral") ~ "Oral",
      str_detect(surg_type, "minor") ~ "Misc minor",
      str_detect(surg_type, "None") ~ "None (healthy)",
      str_detect(surg_type, "abdomin") ~ "Other GI",
      str_detect(surg_type, "Endos") ~ "Endoscopy"),
    country = ifelse(grepl("USA", country), "USA", country),
    country = ifelse(grepl("UK", country), "UK", country),
    country = ifelse(!is.na(non_vh_hdi), str_c(country, "^a^"), country),
    n_analyze = ifelse(refid == 2009, 9, n_analyze),
    study = ifelse(grepl("Marquini", study), str_c(study, "^b^"), study),
    surg = ifelse(surg_nosurg == "nonsurgical", "None", surg)) %>%
  arrange(age, desc(surg_nosurg), design, year, study) %>%
  select(refid, study, n_analyze, centers, country, surg) %>%
  kbl(
    booktabs = T, align = c("llccll"), # format = "latex",
    col.names = c("        ID", "Study", "Analyzed (N)", "Centers", "Country^a^", "Surgery")) %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  column_spec(1, width = "8em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "8em") %>%
  column_spec(4, width = "5em") %>% 
  column_spec(c(5,6), width = "9em") %>% 
  pack_top(., "Adult, Surgical", 1, 7) %>% 
  pack_sub(., "RCT", 1, 6) %>% 
  pack_sub(., "Case Series", 7, 7) %>% 
  pack_top(., "Adult, Non-surgical", 8, 9) %>% 
  pack_sub(., "Crossover", 8, 8) %>% 
  pack_sub(., "Single Arm Trial", 9, 9) %>% 
  pack_top(., "Pediatric, Surgical", 10, 10) %>% 
  pack_sub(., "RCT", 10, 10) %>%
  footnote(
    alphabet = c(
      "Non very high [Human Development Index](https://en.wikipedia.org/wiki/List_of_countries_by_Human_Development_Index) country."),
    general_title = "",
    footnote_as_chunk = FALSE)

table_n <- tab_inc()
```

<br>

<hr style="height:2px;border-width:0;color:gray;background-color:gray">


# **Patient Reported Outcomes**

<a id="gumTab2"></a>

<!-- <manchor id="gumTab2"></manchor> -->

## *Thirst* 

<br/>

<font size = 4> Table `r table_n`. Patient-rated thirst following fasting or gum in randomized controlled trials. </font>

```{r thirstData, echo = FALSE}
gum_thirst <- readxl::read_xlsx("data/GumOutcomes_102820.xlsx", sheet = "thirst", range = "A7:I13") %>% 
  janitor::clean_names()

gum_row <- gum_thirst %>% 
  mutate(row = row_number()) %>% 
  filter(str_detect(tx, "[Gg]um") == TRUE) %>% 
  pull(row)
```

```{r thirstKable, echo = FALSE}
gum_thirst %>% 
  add_row() %>% 
  select(-refid) %>%
  kbl(booktabs = T, align = c("llcccccc"), escape = FALSE,
  col.names = c("        Author", "Treatment", "N", "Time", "Chewing", "Scale^a^", "Median", "IQR")) %>%    
  row_spec(0, bold = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  row_spec(gum_row, color = "brown") %>% 
  add_header_above(c(" " = 3, "Last Chewing" = 1, "Time" = 1, " " = 3), line = FALSE, bold = TRUE) %>% 
  column_spec(1, width = "11em") %>%
  column_spec(2, width = "14em") %>%
  column_spec(3, width = "3em") %>%
  column_spec(4, width = "7em") %>%
  column_spec(5, width = "6em") %>%
  column_spec(6, width = "6em") %>%
  column_spec(7, width = "4em") %>%
  column_spec(8, width = "5em") %>%
  pack_top(., "Adult, Surgical", 1, 6) %>% 
  pack_sub(., "RCT", 1, 6) %>% 
  result_pack("Lower in gum arm, p &lt; .001", 3, padding = 75) %>% 
  result_pack("Difference not detected", 5, padding = 75) %>% 
  result_pack("Less in gum arm, P = .03", 7, padding = 75) %>% 
  footnote(general = "IQR: interquartile range; RCT: randomized controlled trial; NR: not reported; VAS: visual analogue scale.", 
    general_title = "",
    alphabet = c("Arrow (→) indicates least to most thirst.",
                 "1 = less in OR compared to morning; 2 = same as morning; 3 = more than morning."),
    footnote_as_chunk = FALSE)

table_n <- tab_inc()
```

<br/>

<a id="gumTab3"></a>

## *Hunger*  

<br style="line-height: 5px" />

<font size = 4> Table `r table_n`. Patient-rated hunger following fasting or gum. </font>

```{r hungerData, echo = FALSE}
gum_hunger <- readxl::read_xlsx("data/GumOutcomes_102820.xlsx", sheet = "hunger", range = "A8:I12") %>% 
  janitor::clean_names()

gum_row <- gum_hunger %>% 
  mutate(row = row_number()) %>% 
  filter(str_detect(tx, "[Gg]um") == TRUE) %>% 
  pull(row)
```

```{r hungerKable, echo = FALSE}
gum_hunger %>% 
  add_row() %>% 
  select(-refid) %>%
  kbl(booktabs = T, align = c("llcccccc"), escape = FALSE,
  col.names = c("        Author", "Treatment", "N", "Time", "Chewing", "Scale", "Median", "IQR")) %>%    
  row_spec(0, bold = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  row_spec(gum_row, color = "brown") %>% 
  add_header_above(c(" " = 3, "Last Chewing" = 1, "Time" = 1, " " = 3), line = FALSE, bold = TRUE) %>% 
  column_spec(1, width = "11em") %>%
  column_spec(2, width = "14em") %>%
  column_spec(3, width = "3em") %>%
  column_spec(4, width = "7em") %>%
  column_spec(5, width = "6em") %>%
  column_spec(6, width = "5em") %>%
  column_spec(7, width = "5em") %>%
  column_spec(8, width = "5em") %>%
  pack_top(., "Adult, Surgical", 1, 4) %>% 
  pack_sub(., "RCT", 1, 4) %>% 
  result_pack("Difference not detected", 3, padding = 75) %>% 
  result_pack("Difference not detected", 5, padding = 75) %>% 
  footnote(general = "IQR: interquartile range; RCT: randomized controlled trial.", 
    general_title = "",
    alphabet = c("1 = less in OR compared to morning; 2 = same as morning; 3 = more than morning."),
    footnote_as_chunk = FALSE)

table_n <- tab_inc()
```

<br/>

<hr style="height:2px;border-width:0;color:gray;background-color:gray">

# **Clinical Outcomes**

<a id="gumTab4"></a>

## *Residual Gastric Volume*  

<br/>

<font size = 4> Table `r table_n`. Residual gastric volumes at induction in randomized controlled trials of adult patients undergoing surgery according to fasting and gum chewing. </font>

```{r rgvData, echo = FALSE}
## (updated 2020/12/17 20:43) rgvData -------------------------------------
# Hamid 2012
# combine_contin(30, 30, 25, 27, 27, 27)
# Dubin 1994 arms 2, 3
# combine_contin(46, 15, 28, 40, 19, 30)
# arms 2, 3, 4
# combine_contin(61, 61, 30.95082, 31, 22.53658, 22)
# NOTE: 

gum_rgv <- readxl::read_xlsx("data/GumOutcomes_102820.xlsx", sheet = "RGV", range = "A7:L42") %>% 
  filter(!is.na(refid)) %>%
  # delete combined arms for 6038 Dubin 6287 Hamid
  slice(-c(4, 10)) %>%
  mutate(
    rangel = as.numeric(str_split(range, " to ", simplify = TRUE)[, 1]),
    rangeu = as.numeric(str_split(range, " to ", simplify = TRUE)[, 2]),
    iqrl = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 1]),
    iqru = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 2]),
    study = str_replace(study, "\\(", ""),
    study = str_replace(study, "\\)", ""),
    # tx = ifelse(str_detect(tx, "free"), "Gum sugarfree", tx),
    # tx = ifelse(str_detect(tx, "sugared"), "Gum sugared", tx),
    range = str_c("(", round(rangel, 0), "-", round(rangeu, 0), ")"),
    iqr = str_c("(", round(iqrl, 0), "-", round(iqru, 0), ")"),
    iqr = ifelse(refid == 4926, str_c("(", round(iqrl, 2), "-", round(iqru, 2), ")"), iqr),
    mean_sd = paste0(formatC(mean, 2), " (", formatC(sd, 2), ")"),
    mean_sd = str_replace(mean_sd, "\\( ", "\\("),
    mean_sd = str_trim(mean_sd),
    mean_sd = na_if(mean_sd, "NA (NA)"),
    mean_sd = ifelse(refid == 2364, str_c(mean_sd, "^c^"), mean_sd),
    med = ifelse(refid == 4926, str_c(med, "^cde^"), med),
    iqr = ifelse(refid == 4926, str_c(iqr, "^cde^"), iqr),
    tx = ifelse(refid == 8584 & n %in% c(21, 23), str_c(tx, "^b^"), tx)
  )

# smaller GFV as compared with children who chewed sugared and sugarless gum (0.35 [0.2– 0.5] mL/kg versus 0.88 [0.6 –1.4] mL/kg versus 0.69 [0.4 –1.6] mL/kg; P  0.0001). Also, a significantly smaller proportion of subjects who did not chew gum had a GFV of more than 0.8 mL/kg as compared with subjects who chewed sugared and sugarless gum (6.3% versus 66.7% versus 40%; P  0.002).
# merge with design and age

gum_rgv <- left_join(gum_rgv,
  study_char_gum.dat[, c("refid", "year", "age", "design", "surg_nosurg")],
  by = "refid"
) %>%
  arrange(age, desc(surg_nosurg), design, year, study) %>%
  fill(study) %>% 
  mutate(tx = ifelse(grepl("Bouvet", study) & grepl("Gum", tx), "Gum Sugared/Water", tx))

```

```{r rgvKable, echo = FALSE}
## (updated 2020/12/17 20:43) rgvKable ------------------------------------
# cell_spec(gum_rgv[study == "Valencia 2019", 2], color = "black")
# gum rows to highlight

gum_row <- gum_rgv %>% 
  mutate(row = row_number()) %>% 
  filter(str_detect(tx, "[Gg]um") == TRUE) %>% 
  pull(row)

gum_rgv$study <- ifelse(gum_rgv$study == "Valencia 2019", cell_spec(gum_rgv$study, color = "black"), gum_rgv$study)

gum_rgv %>% 
  group_by(refid) %>% 
    mutate(study = ifelse(row_number() == 1, study, "")) %>% 
  ungroup() %>% 
  select(study:time_meas, mean_sd, med, iqr, range) %>%
  mutate(n = ifelse(is.na(n), "NR", n)) %>% 
  kable(booktabs = T, align = c("llcccccccc"), escape = FALSE,
  col.names = c("        Study", "Treatment", "N", "Time (hr)^a^", "Measure", "Time", "Mean (SD)", "Median", "IQR", "Range")) %>%    
  add_header_above(c(" " = 3, " " = 1, "RGV" = 6), line = TRUE, bold = TRUE) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  row_spec(gum_row, color = "brown") %>% 
  column_spec(1, width = "12em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "3em") %>%
  column_spec(4, width = "6em") %>%
  column_spec(5, width = "3em") %>%
  column_spec(6, width = "3em") %>%
  column_spec(7, width = "6em") %>%
  column_spec(8, width = "4em") %>%
  column_spec(9, width = "6em") %>%
  column_spec(10, width = "4em") %>%
  pack_top(., "Adult, Surgical", 1, 16) %>% 
  pack_sub(., "RCT", 1, 16) %>% 
  pack_top(., "Adult, Non-surgical", 17, 19) %>% 
  pack_sub(., "Crossover", 17, 18) %>% 
  result_pack("Difference not detected", 19, padding = 62) %>% 
  pack_sub(., "Single Arm Trial", 19, 19) %>% 
  pack_top(., "Pediatric, Surgical", 20, 22) %>% 
  pack_sub(., "RCT", 20, 22) %>% 
  footnote(general = "RGV: residual gastric volume; SD: standard deviation; IQR: interquartile range; RCT: randomized controlled trial; Bicarb: bicarbonate; Asp: aspiration; US: ultra- sound; NR: not reported.", 
    general_title = "",
    alphabet = c("Before induction.", 
                 "Smokers.",
                 "mL/kg.",
                 "2 hr after ingestion of 250 mL and gum chewing.",
                 "p &lt; 0.001 for comparison with fasting."),
    footnote_as_chunk = FALSE)

table_n <- table_n + 1
```

<br>

<a id="gumFig1"></a>

### Pooled Results 

<br/>

<font size = 4> Figure `r figure_n`. Forest plot for pooled mean difference for residual gastric volume in randomized controlled trials of adult surgical patients --- gum chewing versus fasting. </font>

```{r rgvGumMeta, include = TRUE, echo = FALSE, fig.width = 9.75, fig.height = 5.2, fig.align = "left"}
gum_rgv <- readxl::read_xlsx("data/GumOutcomes_102820.xlsx", range = "A7:L42")

temp_rgv <- gum_rgv %>%
  filter(!is.na(refid)) %>%
  filter(!refid %in% c(4926, 433, 1314)) %>%
  # delete single arms for 6038 Dubin, 6287 Hamid loolipop water
  slice(-c(2, 3, 6:9)) %>%
  mutate(
    rangel = as.numeric(str_split(range, " to ", simplify = TRUE)[, 1]),
    rangeu = as.numeric(str_split(range, " to ", simplify = TRUE)[, 2]),
    iqrl = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 1]),
    iqru = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 2]),
    study = str_replace(study, "\\(", ""),
    study = str_replace(study, "\\)", ""),
    tx = ifelse(str_detect(tx, "free"), "Gum", tx),
    tx = ifelse(str_detect(tx, "ugared"), "Gum", tx)
  ) %>%
  select(refid:n, mean, sd, med, rangel, rangeu, iqrl, iqru) %>%
  mutate(
    mean = ifelse(refid == 2364, mean * 70, mean),
    sd = ifelse(refid == 2364, sd * 70, sd),
    # add subgroup for Soreide and suppl refid
    refid = ifelse(refid == 8584 & n != 30, 85840, refid),
    study = ifelse(refid == 85840 & n == 21, "Soreide 1995", study),
    smoker = ifelse(refid == 85840, "Smokers", "Non-smokers")
  )

rgv_meta.dat <- left_join(temp_rgv[c(1, 3, 5, 7, 9, 11), ], temp_rgv[c(2, 4, 6, 8, 10, 12), ],
  by = c("refid")
)

rgv_meta <- metacont(
  n.e = n.y, n.c = n.x,
  mean.e = mean.y,
  sd.e = sd.y,
  median.e = med.y,
  q1.e = iqrl.y,
  q3.e = iqru.y,
  min.e = rangeu.y,
  max.e = rangel.y,
  mean.c = mean.x,
  sd.c = sd.x,
  median.c = med.x,
  q1.c = iqrl.x,
  q3.c = iqru.x,
  min.c = rangeu.x,
  max.c = rangel.x,
  data = rgv_meta.dat,
  method.tau = "REML",
  subgroup = smoker.x,
  studlab = study.x
)

forest(rgv_meta,
  fixed = FALSE,
  lab.e = "Gum         ",
  lab.c = "Fasting     ",
  digits = 1,
  digits.sd = 1,
  digits.tau2 = 1,
  # prediction = TRUE,
  rightlabs = c("MD", "95% CI"),
  xlim = c(-20, 30),
  fs.xlab = 11,
  print.subgroup.name = FALSE,
  rightcols = c("effect", "ci"),
  xlab = "RGV (mL)   \n Favors Gum                    Favors Fasting"
)

figure_n <- fig_inc()
```

<font size = 2> MD: mean difference;RGV: residual gastric volume. <br>Gum arms combined for Dubin 1994 and Hamid 2012. Goudra 2015 estimates assume a 70 kg patient for pooling.</font>

<br/>

<a id="gumTab5"></a>

## *Gastric pH* 

<br/>

```{r phData, echo = FALSE}
## (updated 2020/12/17 20:49) phData --------------------------------------
# Hamid 2012
# combine_contin(30, 30, 3, 5, 5, 5)
temp_ph <- readxl::read_xlsx("data/GumOutcomes_102820.xlsx", sheet = "pH", range = "A8:L39") %>% 
  janitor::clean_names()

gum_ph <- temp_ph %>%
  filter(!is.na(refid)) %>%
  # delete combined arms for 6038 Dubin 6287 Hamid
  slice(-c(4, 10)) %>%
  mutate(
    rangel = as.numeric(str_split(range, " to ", simplify = TRUE)[, 1]),
    rangeu = as.numeric(str_split(range, " to ", simplify = TRUE)[, 2]),
    iqrl = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 1]),
    iqru = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 2]),
    study = str_replace(study, "\\(", ""),
    study = str_replace(study, "\\)", ""),
    range = str_c("(", formatC(rangel, format = "f", 1), "-", round(rangeu, 1), ")"),
    iqr = str_c("(", round(iqrl, 1), "-", round(iqru, 1), ")"),
    mean_sd = paste0(round(as.numeric(mean), 1), " (", round(sd, 1), ")"),
    mean_sd = ifelse(refid == 8584, paste0(as.numeric(mean), " (", round_1(sd), ")"), mean_sd),
    mean_sd = ifelse(is.na(sd), round(as.numeric(mean), 1), mean_sd),
    mean_sd = str_replace(mean_sd, "\\( ", "\\("),
    mean_sd = str_trim(mean_sd),
    mean_sd = na_if(mean_sd, "NA (NA)"))

# merge with design and age
gum_ph <- left_join(gum_ph,
  study_char_gum.dat[, c("refid", "year", "age", "design", "surg_nosurg")],
  by = "refid") %>%
  filter(refid != 433) %>%  # Valencia did not report
  arrange(age, desc(surg_nosurg), design, year, study) %>% 
  rename(tx = intervention) %>% 
  mutate(tx = ifelse(refid == 8584 & n %in% c(21, 23), str_c(tx, "^b^"), tx))

```

<font size = 4> Table `r table_n`. Gastric pH at induction in randomized controlled trials of patients undergoing surgery according to fasting, liquid and volume. </font>

```{r phKable, echo = FALSE}
## (updated 2020/12/17 20:50) phKable -------------------------------------
# cell_spec(gum_rgv[study == "Valencia 2019", 2], color = "black")
gum_row <- gum_ph %>% 
  mutate(row = row_number()) %>% 
  filter(str_detect(tx, "[Gg]um") == TRUE) %>% 
  pull(row)

gum_ph %>% 
  select(study:time, time_meas, mean_sd, med, range, iqr) %>%
  mutate(n = ifelse(is.na(n), "NR", n)) %>% 
  kable(booktabs = T, align = c("llcccllll"), escape = FALSE,
  col.names = c("          Author", "Treatment", "N", "Time (hr)^a^", "Measured", "Mean (SD)", "Median", "Range", "IQR")) %>%    
  add_header_above(c(" " = 3, " " = 1, "Time" = 1, " " = 4), bold = TRUE, line = FALSE) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  row_spec(gum_row, color = "brown") %>% 
  column_spec(1, width = "12em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "3em") %>%
  column_spec(4, width = "7em") %>%
  column_spec(5, width = "7em") %>%
  column_spec(6, width = "7em") %>%
  column_spec(7, width = "5em") %>%
  column_spec(8, width = "5em") %>%
  column_spec(9, width = "5em") %>%
  pack_top(., "Adult, Surgical", 1, 16) %>% 
  pack_sub(., "RCT", 1, 16) %>% 
  pack_top(., "Pediatric, Surgical", 17, 19) %>% 
  pack_sub(., "RCT", 17, 19) %>% 
  footnote(general = "SD: standard deviation; IQR: interquartile range; RCT: randomized controlled trial; Bicarb: bicarbonate; Asp: aspiration; NR: not reported.", 
    general_title = "",
    alphabet = c("Before induction.",
                 "Smokers."),
    footnote_as_chunk = FALSE)

table_n <- tab_inc()
```

<br/>

<a id="gumFig2"></a>

### Pooled Results 

<br/>

<font size = 4> Figure `r figure_n`. Forest plot for pooled mean difference for pH in randomized controlled trials of adult surgical patients --- gum chewing versus fasting. </font>

```{r phGumMeta, include = TRUE, echo = FALSE, fig.width = 9.75, fig.height = 5.5, fig.align = "top"}
temp_ph <- readxl::read_xlsx("data/GumOutcomes_102820.xlsx", sheet = "pH", range = "A8:L39") %>% 
  janitor::clean_names()

temp_ph <- temp_ph %>%
  filter(!is.na(refid)) %>%
  # keep combined arms for 6038 Dubin 6287 Hamid
  slice(-c(2, 3, 6:9)) %>%
  filter(refid != 4926) %>% # Schoenfeld no n per arm
  mutate(
    rangel = as.numeric(str_split(range, " to ", simplify = TRUE)[, 1]),
    rangeu = as.numeric(str_split(range, " to ", simplify = TRUE)[, 2]),
    iqrl = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 1]),
    iqru = as.numeric(str_split(iqr, " to ", simplify = TRUE)[, 2]),
    study = str_replace(study, "\\(", ""),
    study = str_replace(study, "\\)", ""),
    # tx = ifelse(str_detect(tx, "free"), "Gum sugarfree", tx),
    # tx = ifelse(str_detect(tx, "sugared"), "Gum sugared", tx),
    range = str_c("(", formatC(rangel, format = "f", 1), "-", round(rangeu, 1), ")"),
    iqr = str_c("(", round(iqrl, 1), "-", round(iqru, 1), ")"),
    mean_sd = paste0(round(as.numeric(mean), 1), " (", round(sd, 1), ")"),
    mean_sd = str_replace(mean_sd, "\\( ", "\\("),
    mean_sd = str_trim(mean_sd),
    mean_sd = na_if(mean_sd, "NA (NA)"),
    # add subgroup for Soreide and suppl refid
    refid = ifelse(refid == 8584 & n != 30, 85840, refid),
    study = ifelse(refid == 85840 & n == 21, "Soreide 1995", study),
    smoker = ifelse(refid == 85840, "Smokers", "Non-smokers")
  )

# merge with design and age
gum_ph <- left_join(temp_ph,
  study_char_gum.dat[, c("refid", "year")], by = "refid") %>%
  rename(tx = intervention) %>% 
  filter(refid != 433) %>%  # Valencia did not report
  select(refid, study, tx, n, mean, sd, med, rangel, rangeu, iqrl, iqru, year, smoker) %>% 
  mutate(n = as.numeric(n), mean = as.numeric(mean))

ph_meta.dat <- left_join(gum_ph[c(1,3,5,7,9,11), ], gum_ph[c(2,4,6,8,10,12), ], 
                          by = c("refid"))
  
ph_meta <- metacont(
  n.e = n.y, n.c = n.x,
  mean.e = mean.y,
  sd.e = sd.y,
  median.e = med.y,
  q1.e = iqrl.y,
  q3.e = iqru.y,
  min.e = rangeu.y,
  max.e = rangel.y,
  mean.c = mean.x,
  sd.c = sd.x,
  median.c = med.x,
  q1.c = iqrl.x,
  q3.c = iqru.x,
  min.c = rangeu.x,
  max.c = rangel.x,
  data = ph_meta.dat,
  method.tau = "REML",
  subgroup = smoker.x,
  studlab = study.x
)

forest(ph_meta,
  fixed = FALSE,
  lab.e = "Gum         ",
  lab.c = "Fasting     ",
  digits = 1,
  digits.sd = 1,
  digits.tau2 = 1,
  prediction = FALSE,
  rightlabs = c("MD", "95% CI"),
  rightcols = c("effect", "ci"),
  xlim = c(-2, 3),
  fs.xlab = 11,
  print.subgroup.name = FALSE,
  subgroup = TRUE,
  xlab = "pH\n Favors Fasting                Favors Gum"
)

figure_n <- fig_inc()
```

<font size = 2> MD: mean difference. <br>Gum arms combined for Dubin 1994 and Hamid 2012 .</font>

<br>

# **Study/Participant Detail**

## *Study Characteristics*

<br/>

<font size = 4> Table `r table_n`. Characteristics of included studies examining chewing gum. </font>
```{r createStudyCharTable, include = FALSE}
## (updated 2021/05/16 10:27) study_char_table ----------------------------
study_char_table <- 
  study_char_gum.dat %>%
  mutate(dates = str_c(date_start, "\U2012", date_end),
    country = countrycode::countryname(country, destination = "iso3c"),
    low_r = noquote(if_else(non_vh_hdi == "yes", "\U2022", "", "")),
    n = n_enrolled,
    n_anal = n_analyze,
    pilot = noquote(if_else(pilot == "yes", "\U2022", "", "")),
    setting = case_when(!is.na(hospital) ~ "Hosp",
      !is.na(ambulatory) ~ "Amb",
      !is.na(oth_setting) ~"Other"),
    gen = noquote(if_else(!is.na(general), "\U2022", "", "")),
    reg = noquote(if_else(!is.na(regional), "\U2022", "", "")),
    sed = noquote(if_else(!is.na(sedation), "\U2022", "", "")),
    registered = noquote(if_else(registered == "yes", "\U2022", "", "")),
    surg = case_when(str_detect(surg_type, "[Gg]yn") ~ "Gyn",
      str_detect(surg_type, "ariety") ~ "Various",
      str_detect(surg_type, "cholecystectomy") ~ "Cholecystectomy",
      str_detect(surg_type, "[Gg]ast|Bowel") ~ "Other GI",
      str_detect(surg_type, "not appli") ~ "None (healthy)",
      str_detect(surg_type, "oral") ~ "Oral",
      str_detect(surg_type, "minor") ~ "Misc minor",
      str_detect(surg_type, "None") ~ "None (healthy)",
      str_detect(surg_type, "Elective") ~ "Elective",
      str_detect(surg_type, "detail") ~ "No detail",
      str_detect(surg_type, "no descr") ~ "No detail",
      str_detect(surg_type, "oscopy") ~ "Endoscopy",
      str_detect(surg_type, "Endso") ~ "Endoscopy"),
    surg = ifelse(refid == 2009, "No detail", surg)) %>%
  mutate(n = ifelse(refid == 2009, 9, n)) %>% 
  # mutate(n = foot_note(353, n, "a")) %>% 
  # mutate(surg = ifelse(refid == 353, str_c(surg, "$^a$"), surg)) %>% # footnote
  arrange(age, desc(surg_nosurg), design, year, study) %>%
  select(study, dates, country, n, pilot, setting, gen, reg, sed, surg, registered, design)

```

```{r studyCharKable, echo = FALSE}
## (updated 2021/05/16 10:42) studyCharKable ------------------------------
study_char_table %>%
  select(-design) %>%
  kable(
    booktabs = T, align = c("lcccccccclc"),
    # format = "latex",
    col.names = c(
      "        Study", "Dates", "Country", " (N\\) ", "Pilot",
      "Setting", "Gen", "Reg", "Sed", "Type of Surgery", "Registered"
    )
  ) %>%
  kable_classic(full_width = TRUE, html_font = "Source Sans Pro", position = "left", "hover") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  column_spec(1, width = "15em") %>%
  # column_spec(2, width = "2em") %>%
  column_spec(3, width = "7em") %>%
  add_header_above(c(" " = 3, "Enrolled" = 1, " " = 2, "Anesthetic" = 3, " " = 2)) %>%
  pack_top(., "Adult, Surgical", 1, 7) %>%
  pack_sub(., "RCT", 1, 6) %>%
  pack_sub(., "Case Series", 7, 7) %>%
  pack_top(., "Adult, Non-surgical", 8, 9) %>%
  pack_sub(., "Crossover", 8, 8) %>%
  pack_sub(., "Single Arm Trial", 9, 9) %>%
  pack_top(., "Pediatric, Surgical", 10, 10) %>%
  pack_sub(., "RCT", 10, 10) %>%
  footnote(
    general = "Gen: general; Reg: regional; Sed: sedation; Hosp: hospital; Amb: ambulatory.",
    general_title = "",
    footnote_as_chunk = T
  ) # %>%
# kable_styling(position = "left", "condensed")

table_n <- tab_inc()
```

<br/><br/>

## *Sample Characteristics*

<br/>

<font size = 4> Table `r table_n`. Characteristics of patients in included studies. </font>
```{r createPtCharTable, include = FALSE}
## (updated 2021/05/16 10:43) createPtCharTable ---------------------------
pt_char_table <- study_arm_gum.dat %>%
  group_by(refid) %>%
  summarize(
    N = sum(arm_n),
    wgt = arm_n / N,
    asai_all = round(sum(asai * wgt, na.rm = TRUE)),
    asaii_all = round(sum(asaii * wgt, na.rm = TRUE)),
    asaiii_all = round(sum(asaiii * wgt, na.rm = TRUE)),
    asai_ii_all = round(sum(asai_ii * wgt, na.rm = TRUE)),
    age_mn_all = round(sum(age_mean * wgt, na.rm = TRUE)),
    age_md_all = round(sum(age_med * wgt, na.rm = TRUE)),
    sex_all = round(sum(female * wgt, na.rm = TRUE)),
    white_all = round(sum(white * wgt, na.rm = TRUE)),
    black_all = round(sum(black * wgt, na.rm = TRUE)),
    asian_all = round(sum(asian * wgt, na.rm = TRUE)),
    bmi_mn_all = round(sum(bmi_mean * wgt, na.rm = TRUE)),
    bmi_md_all = round(sum(bmi_med * wgt, na.rm = TRUE)),
    dm_mn_all = round(sum(dm * wgt, na.rm = TRUE)),
    neuro_all = round(sum(neuro * wgt, na.rm = TRUE))
  ) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  mutate(
    N = ifelse(refid == 4926, 46, N),
    age_mn_all = ifelse(refid == 4926, "5-17^a^", age_mn_all),
    age_md_all = ifelse(refid == 4926, " ", age_md_all)
  )
  
pt_char_table <- left_join(pt_char_table,
  study_char.dat[, c("refid", "study", "design", "year", "age", "surg_nosurg")],
  by = "refid") %>%
  arrange(age, desc(surg_nosurg), design, year, study) %>%
  select(design, study, year, everything())

pt_char_table <- pt_char_table %>% 
  select(study, N, asai_all, asaii_all, asai_ii_all, asaiii_all, sex_all, age_mn_all, 
  age_md_all, white_all, black_all, asian_all, bmi_mn_all, bmi_md_all, dm_mn_all)
  
```

```{r ptCharKable, echo = FALSE}
## (updated 2021/05/16 10:45) ptCharKable ---------------------------------
pt_char_table %>% 
  mutate_at(vars(asai_all:dm_mn_all), na_if, 0) %>% 
  kable(booktabs = T, align = c("lcrrrcccccccccc"),
    col.names = c("\\ \\ \\ \\ \\ \\ \\ Study", "Analyzed", "I", "II", "I-II", "III", "(%)",
      "Mean", "Med", "White", "Black", "Asian", "Mean", "Med", "DM (%)")) %>%
  add_header_above(c(" " = 1, "N" = 1, "ASA Classs (%)" = 4, "Female" = 1, "Age" = 2,
    "Race (%)" = 3, "BMI (kg/m$^2$)" = 2, " " = 1), line = TRUE) %>%
  kable_classic(full_width = TRUE, html_font = "Source Sans Pro", position = "left", "hover") %>% 
  kable_styling(bootstrap_options = c("striped", "hover","condensed")) %>% 
  column_spec(1, width = "15em") %>%
  # column_spec(2, width = "4em") %>%
  column_spec(3, width = "3em") %>%
  pack_top(., "Adult, Surgical", 1, 7) %>% 
  pack_sub(., "RCT", 1, 6) %>% 
  pack_sub(., "Case Series", 7, 7) %>% 
  pack_top(., "Adult, Non-surgical", 8, 9) %>% 
  pack_sub(., "Crossover", 8, 8) %>% 
  pack_sub(., "Case Series", 9, 9) %>% 
  pack_top(., "Pediatric, Surgical", 10, 10) %>% 
  pack_sub(., "RCT", 10, 10) %>% 
  footnote(
    general = "Med: median",
    alphabet = c("Range"),
    general_title = "",
    footnote_as_chunk = T,
  )

table_n <- tab_inc()
```

<br/>

## *Fasting and Gum Timing*

<br/>

<font size = 4> Table `r table_n`. Fasting times and gum chewing protocols according to study arm. </font>
```{r createFastingTable, include = FALSE}
## (updated 2021/05/16 10:45) createFastingTable --------------------------
fasting_table <- study_arm.dat %>% 
  filter(refid %in% gum_refids$refid) %>% 
  mutate(arm_tx = factor(arm_tx,
    levels = c("fasting", "water", "other", "gum_sugared", "gum_sugarfree"))) %>% 
  mutate(tx = str_replace(arm_tx, "_", " "),
    tx = str_to_title(tx), 
    tx = if_else(tx == "Cho", "CHO", tx, ""),
    tx = ifelse(refid == 6287 & tx == "Other", "Lollipop", tx),
    tx = ifelse(refid == 2009, "Gum (any)", tx)) %>% 
  select(refid, tx, arm_n, fastsolid, fastliquid, type_of_liquid, timeingest1, amtingest1, timeingest2, amtingest2, age, surg_nosurg, design, arm_tx, year, study) %>% 
  arrange(age, desc(surg_nosurg), design, year, study, arm_tx) %>% 
  select(refid, study, year, design, everything())

#fasting_table <- left_join(fasting_table, study_names, by = "refid") %>% 
  #select(refid, study, year, design, everything())
  #arrange(age, desc(surg_nosurg), design, year, study) %>%
  #group_by(study)

temp <- read_csv("data/fasting_table_add_102720.csv")[,c(1,5,6,14,15)]

fasting_table <- left_join(fasting_table, temp, by = c("refid", "tx", "arm_n"))

rm(temp)

# write_csv(fasting_table, "/Users/mgrant/Documents/_projects01/asa/_preopFasting/5_analysis/_r/_data/fasting_table_add.csv")

fasting_table <- fasting_table %>%
  # mutate(tx = factor(tx, levels = c("Fasting", "Water", "Gum Sugared", "Gum Sugarfree", "Gum Nicotine", "Gum (any)", "Lollipop"))) %>% 
  mutate(timeingest1 = ifelse(timeingest1 == 99, 12, timeingest1)) %>% 
  mutate(amtingest1  = na_if(amtingest1, 0)) %>% 
  mutate(timeingest1 = round(timeingest1, 0)) %>% 
  mutate(timeingest2 = round(timeingest2, 0)) %>% 
  arrange(age, desc(surg_nosurg), design, year, study) %>% 
  group_by(study) %>% 
  # mutate(design = if_else(row_number() == 1, design, "")) %>% 
  mutate(year = ifelse(row_number() == 1, year, NA ),
         study = if_else(row_number() == 1, study, "")) %>% 
  ungroup() %>% 
  mutate(study = ifelse(study %in% c("Valencia 2019", "Shanmugam 2016"), cell_spec(study, color = "black"), study))

```

```{r fastingKable, echo = FALSE}
gum_row <- fasting_table  %>% 
  mutate(row = row_number()) %>% 
  filter(str_detect(tx, "[Gg]um") == TRUE) %>% 
  pull(row)

# fasting_table$study <- ifelse(fasting_table$study == "Valencia 2019", cell_spec(fasting_table$study, color = "black"), fasting_table$study)

fasting_table %>%
  select(study, tx, arm_n, fastsolid, fastliquid, start_gum_oth, stop_gum_oth) %>% 
  add_column(temp = NA) %>% 
  select(study:fastliquid, temp, everything()) %>% 
  kbl(booktabs = T, align = c("llccccll"), escape = FALSE,
      col.names = c("        Study", "Comparator", "N", "Solids", "Liquids", " ",  "Chewing or Time", "Gum Chewing")) %>% 
  add_header_above(c(" " = 3, "   Prescribed Fasting (hr)" = 2, " " = 1,  "   Start Gum" = 1, "   Stop" = 1), align = "l", underline = FALSE) %>% 
  kable_classic(full_width = TRUE, html_font = "Source Sans Pro", position = "left", "hover") %>% 
  kable_styling(bootstrap_options = c("striped", "hover","condensed")) %>% 
  row_spec(gum_row, color = "brown") %>% 
  column_spec(1, width = "11em") %>% 
  column_spec(2, width = "8em") %>% 
  column_spec(c(4,5), width = "6em") %>%
  column_spec(6, width = "1em") %>% 
  pack_top(., "Adult, Surgical", 1, 20) %>% 
  pack_sub(., "RCT", 1, 19) %>% 
  pack_sub(., "Case Series", 20, 20) %>% 
  pack_top(., "Adult, Non-surgical", 21, 23) %>% 
  pack_sub(., "Crossover", 21, 22) %>% 
  pack_sub(., "Single Arm Trial", 23, 23) %>% 
  pack_top(., "Pediatric, Surgical", 24, 26) %>% 
  pack_sub(., "RCT", 24, 26) %>%  
  footnote(alphabet = c("Hours prior to surgery."),
          alphabet_title = " ",
          general_title = "",
          footnote_as_chunk = T)

  table_n <- tab_inc()
```

<br><br>

## *Comparator Detail*

<br>

<font size = 4> Table `r table_n`. Comparator details.</font>

```{r createComparatorDetailTable, echo = FALSE}
comparator_detail_table <- study_arm.dat %>% 
  filter(refid %in% gum_refids$refid) %>% 
  mutate(arm_tx = factor(arm_tx,
    levels = c("fasting", "water", "other", "gum_sugared", "gum_sugarfree"))) %>% 
  mutate(tx = str_replace(arm_tx, "_", " "),
         tx = str_replace(tx, "sugarless", "sugarfree"),
         tx = str_replace(tx, "Sugarless", "Sugarfree"),
         tx = str_to_title(tx), 
         tx = if_else(tx == "Cho", "CHO", tx, ""),
         tx = ifelse(refid == 6287 & tx == "Other", "Lollipop", tx),
         tx = ifelse(refid == 2009, "Gum (any)", tx),
         tx_detail = firstup(tx_detail), 
         tx = factor(tx, 
           levels = c("Fasting", "Water", "Lollipop", "Gum Sugarfree", "Gum Sugared", "Gum (any)"))) %>% 
  arrange(age, desc(surg_nosurg), design, year, study, arm_tx) %>% 
  # select(refid, arm_n, tx, type_of_liquid, tx_detail, age, surg_nosurg, design, year, study) %>% 
  select(study, arm_n, tx, tx_detail) %>% 
  group_by(study) %>% 
  mutate(study = if_else(row_number() == 1, study, "")) %>% 
  ungroup() %>% 
  mutate(study = ifelse(study %in% c("Valencia 2019", "Shanmugam 2016"), cell_spec(study, color = "black"), study))

```

```{r createComparatorDetailKable, echo = FALSE}
gum_row <- comparator_detail_table  %>%
  mutate(row = row_number()) %>% 
  filter(str_detect(tx, "[Gg]um") == TRUE) %>% 
  pull(row)

comparator_detail_table %>%
  kbl(
    booktabs = T, align = c("llll"), escape = FALSE,
    col.names = c("        Author", "N", "Comparator", "Detail")
  ) %>%
  kable_classic(full_width = TRUE, html_font = "Source Sans Pro", position = "left", "hover") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(gum_row, color = "brown") %>%
  column_spec(1, width = "15em") %>%
  column_spec(2, width = "5em") %>%
  column_spec(3, width = "8em") %>%
  pack_top(., "Adult, Surgical", 1, 20) %>%
  pack_sub(., "RCT", 1, 19) %>%
  pack_sub(., "Case Series", 20, 20) %>%
  pack_top(., "Adult, Non-surgical", 21, 23) %>%
  pack_sub(., "Crossover", 21, 22) %>%
  pack_sub(., "Single Arm Trial", 23, 23) %>%
  pack_top(., "Pediatric, Surgical", 24, 26) %>%
  pack_sub(., "RCT", 24, 26)

table_n <- tab_inc()
```

<br>

## *Funding*

<br>

<font size = 4> Table `r table_n`. Reported funding sources. </font>
```{r createFundingTable, include = FALSE}
study_fund_table <- study_char_gum.dat %>%
  mutate(
    # sponsor_desc = str_to_title(sponsor_desc),
    registered = noquote(if_else(registered == "yes", "\U2022", "", "")),
    public = noquote(if_else(funding == "public" | funding == "pub_indus", "\U2022", "", "")),
    industry = noquote(if_else(funding == "pub_indus", "\U2022", "", "")),
    none = noquote(if_else(funding == "none", "\U2022", "", "")),
    not_reported = noquote(if_else(funding == "NR", "\U2022", "", ""))) %>%
  arrange(age, desc(surg_nosurg), design, year, study) %>%
  select(study, design, registered, public, industry, none, not_reported, sponsor_desc)

```

```{r fundingKable, echo = FALSE}
study_fund_table %>% 
  select(-design, -registered) %>% 
  kbl(booktabs = T, align = c("lccccl"),
      col.names = c("\\ \\ \\ \\ \\ \\ \\ Author", "Public", "Industry", "None", 
                    "Reported", "Funding description")) %>%
  add_header_above(c(" " = 4, "Not" = 1, " " = 1), line = FALSE) %>% 
  add_header_above(c(" " = 1, "Reported Funding Source(s)" = 4, " " = 1), line = TRUE) %>% 
  kable_classic(full_width = FALSE, html_font = "Source Sans Pro", position = "left", "hover") %>% 
  kable_styling(bootstrap_options = c("striped", "hover","condensed"), position = "left") %>% 
  column_spec(1, width = "11em") %>% 
  column_spec(c(2,3,4,5), width = "5em") %>% 
  column_spec(6, width = "11em") 

table_n <- tab_inc()
```

<br>

# **Risk of Bias** 

### *Patient Reported Outcomes*

<br>

<font size = 4> Figure `r figure_n`. Summary risk of bias appraisal for randomized controlled trials including patient reported outcomes. </font>

<img src="assets/gum_pro_rob_summary.svg" style="width:600px;" align="left"/>

<br clear="all" />

```{r}
figure_n <- fig_inc()
```

<font size = 4> Figure `r figure_n`. Individual study risk of bias appraisal for randomized controlled trials including patient reported outcomes. </font>

<img src="assets/gum_pro_rob_traffic.svg" style="width:600px;" align="left"/>

<br clear="all" />

```{r}
figure_n <- fig_inc()
```

<br>

### *Clinical Outcomes*

<br>

<font size = 4> Figure `r figure_n`. Summary risk of bias appraisal for randomized controlled trials including clinical outcomes. </font>

<img src="assets/gum_clin_rob_summary.svg" style="width:600px;" align="left"/>

<br clear="all" />

```{r}
figure_n <- fig_inc()
```

<font size = 4> Figure `r figure_n`. Individual study risk of bias appraisal for randomized controlled trials for clinical outcomes. </font>

<img src="assets/gum_clin_rob_traffic.svg" style="width:600px;" align="left"/>

```{r}
figure_n <- fig_inc()
```
<br clear="all" />

<hr style="height:2px;border-width:0;color:gray;background-color:gray">

<br>




# References 

<br>




