---
title: "Network Meta-Analyses — Carbohydrates"
date: "`r format(Sys.time(), '%d %B, %Y (%H:%M)')`"
output:
  html_document:
    font: Source Sans Pro
    theme: cosmo
    anchor_sections: TRUE
    toc_float: 
      collapsed: true
      smooth_scroll: TRUE
    toc_depth: 3
    linkcolor: blue
  latex_engine: xelatex
  pdf_document:
    toc: yes
    toc_depth: '3'
bibliography: cho.bib
csl: jama.csl
link-citations: yes
workflowr:
  suppress_report: false
# nocite: '@*'
---

#### 

```{=html}
<style type="text/css">
  body{
  font-size: 9.5pt;
  font-family: Source Sans Pro;
}

.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
  color: #090909;
  background-color: #0909093d;
}

.lightable-classic.lightable-hover tbody tr:hover {
    background-color: #c5d4e4;
}
.btn-xs, .btn-group-xs>.btn { 
display: none; 
}

.btn-default {
display: none !important; 
}
</style>
```

```{r setupData, include = FALSE}
## * (updated 2021/01/06 11:34) preliminaries -------------------------------
library(multinma)
library(BUGSnet)
# functions
now <- function() {str_replace_all(format(Sys.time(), "%m %d %y_%H-%M"), " ", "_")}
m_table <- function(data, x){
  group_by(data, arm_tx) %>%
    summarise(tx_count = n())
}
# replace_mg <- function(a, b) {if_else(is.na(a) & !is.na(b), b, a)}
tab_inc <- function() {
  table_n <- table_n + 1
  table_n}
fig_inc <- function() {
  figure_n <- figure_n + 1
  figure_n}
table_n <- 1
figure_n <- 1

## * get current files; select cho refids; directory must have unique ####
source("code/readFiles_120220_wfr.R")

# study char cho refs 
study_char_cho.dat <- study_char.dat %>% 
  filter(refid %in% cho_refids)

# study arm protein refs 
study_arm_cho.dat <- study_arm.dat %>%
  filter(refid %in% cho_refids) %>%
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>%
  # remove duplicate arms tewari 2019 (refid 716)
  filter(!(refid == 716 & arm_n == 8)) %>%
  mutate(
    study = ifelse(refid == 2564 & grepl("colorectal", notes_arm), paste(study, "(CR)"), study),
    study = ifelse(refid == 2564 & grepl("cholecystectomy", notes_arm), paste(study, "(CH)"), study),
    study = ifelse(refid == 1926 & grepl("general", notes_arm), paste(study, "(gen)"), study),
    study = ifelse(refid == 1926 & grepl("epidural", notes_arm), paste(study, "(epi)"), study)
  ) %>%
  mutate(tx = factor(arm_tx,
    levels = c(
      "fasting",
      "water",
      "placebo",
      "other_clear",
      "milk",
      "prot",
      "prot_simp",
      "prot_comp",
      "cho_simp",
      "cho_comp"
    )
  )) %>%
  mutate(
    tx = fct_recode(tx,
      Fasting = "fasting",
      Water = "water",
      Placebo = "placebo",
      Clear = "other_clear",
      Milk = "milk",
      "Prot" = "prot",
      "CHO/Comp/Prot" = "prot_comp",
      "CHO/Simp/Prot" = "prot_simp",
      "CHO/Simp" = "cho_simp",
      "CHO/Comp" = "cho_comp",
    ),
    arm_tx = tx
  ) %>%
  select(-tx) 

# tewari 2019
study_arm_cho_tewari.dat <- study_arm.dat %>% 
  filter(refid == 716)

## * dichot factor treatment arm outcome type for ordering (updated 2020/12/12 12:04) ####
# save dichot.dat 
dichot_cho.dat <- dichot.dat %>% 
  filter(refid %in% cho_refids) %>% 
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>% 
  # factor arm_tx levels for cho studies
  mutate(tx = factor(arm_tx,
  levels = c(
    "fasting",
    "water",
    "placebo",
    "other_clear",
    "milk",
    "prot",
    "prot_simp",
    "prot_comp",
    "cho_simp",
    "cho_comp"))) %>%
  mutate(tx = fct_recode(tx,
    Fasting = "fasting",
    Water = "water",
    Placebo = "placebo",
    Clear = "other_clear",
    Milk = "milk",
    "Prot" = "prot",
    "CHO/Simp/Prot" = "prot_simp",
    "CHO/Comp/Prot" = "prot_comp",
    "CHO/Simp" = "cho_simp",
    "CHO/Comp" = "cho_comp",),
    arm_tx = tx) %>%
  select(-tx)

dichot_cho.dat %>% tabyl(arm_tx)

## * contin factor treatment arm outcome type for ordering (updated 2020/12/12 12:04) ####
contin_cho.dat <- contin.dat %>% 
  filter(refid %in% cho_refids) %>% 
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>%
  # factor arm_tx levels for cho studies
  mutate(tx = factor(arm_tx,
  levels = c(
    "fasting",
    "water",
    "placebo",
    "other_clear",
    "milk",
    "prot",
    "prot_simp",
    "prot_comp",
    "cho_simp",
    "cho_comp"))) %>%
  mutate(tx = fct_recode(tx,
    Fasting = "fasting",
    Water = "water",
    Placebo = "placebo",
    Clear = "other_clear",
    Milk = "milk",
    "Prot" = "prot",
    "CHO/Simp/Prot" = "prot_simp",
    "CHO/Comp/Prot" = "prot_comp",
    "CHO/Simp" = "cho_simp",
    "CHO/Comp" = "cho_comp",),
    arm_tx = tx) %>%
  select(-tx)

contin_cho.dat %>% tabyl(arm_tx)

## * likert factor treatment arm outcome type for ordering (updated 2020/12/12 12:04) ####
likert_cho.dat <- likert.dat %>% 
  filter(refid %in% cho_refids) %>% 
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>%
  # factor arm_tx levels for cho studies
  mutate(tx = factor(arm_tx,
  levels = c(
    "fasting",
    "water",
    "placebo",
    "other_clear",
    "milk",
    "prot",
    "prot_simp",
    "prot_comp",
    "cho_simp",
    "cho_comp"))) %>%
  mutate(tx = fct_recode(tx,
    Fasting = "fasting",
    Water = "water",
    Placebo = "placebo",
    Clear = "other_clear",
    Milk = "milk",
    "Prot" = "prot",
    "CHO/Simp/Prot" = "prot_simp",
    "CHO/Comp/Prot" = "prot_comp",
    "CHO/Simp" = "cho_simp",
    "CHO/Comp" = "cho_comp",),
    arm_tx = tx) %>%
  select(-tx)

likert_cho.dat %>% tabyl(arm_tx)

## * divide Sada 2014 and Celiksular 2016 into 2 studies ------------------
likert_cho.dat <- likert_cho.dat %>%
  mutate(
    study = ifelse(study == "Sada 2014" & grepl("cholecystectomy", out_l_notes), "Sada 2014 (CH)", study),
    study = ifelse(study == "Sada 2014" & !grepl("cholecystectomy", out_l_notes), "Sada 2014 (CR)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("general", out_l_notes), "Celiksular 2016 (G)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("epidural", out_l_notes), "Celiksular 2016 (E)", study)
  )

contin_cho.dat <- contin_cho.dat %>%
  mutate(
    study = ifelse(study == "Sada 2014" & grepl("cholecystectomy", notes_continuous), "Sada 2014 (CH)", study),
    study = ifelse(study == "Sada 2014" & !grepl("cholecystectomy", notes_continuous), "Sada 2014 (CR)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("general", notes_continuous), "Celiksular 2016 (G)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("epidural", notes_continuous), "Celiksular 2016 (E)", study)
  )

dichot_cho.dat <- dichot_cho.dat %>%
  mutate(
    study = ifelse(study == "Sada 2014" & grepl("cholecystectomy", out_dichot_notes), "Sada 2014 (CH)", study),
    study = ifelse(study == "Sada 2014" & !grepl("cholecystectomy", out_dichot_notes), "Sada 2014 (CR)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("general", out_dichot_notes), "Celiksular 2016 (G)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("epidural", out_dichot_notes), "Celiksular 2016 (E)", study)
  )


```

```{r rgvData, include = TRUE}
## base arm_tx levels ####
# "Fasting"      
# "Water"        
# "Placebo"      
# "Clear"        
# "Milk"         
# "Prot"         
# "CHO/Simp/Prot"
# "CHO/Comp/Prot"
# "CHO/Simp"     
# "CHO/Comp"

## (updated 2021/05/15 07:21) rgv_meta.dat --------------------------------
rgv_meta.dat <- contin_cho.dat %>%
  arrange(design, year, study) %>%
  filter(refid %in% cho_refids, surg_nosurg == "surgical") %>%
  filter(!is.na(rgv_us) | !is.na(rgv_imag) | !is.na(rgv_asp)) %>%
  filter(design == "rct", age == "Adult") %>%
  select(refid, study, year, design, arm_n, arm_tx, rgv_mean_1, rgv_sd_1, rgv_median_1:rgv_iqr_up_1, hr_gt6:hr_2) %>%
  # Yilmaz 2013 estimate sd smd.bwgrpS(21.85, 19.15, 20, 20, .77),
  # Yildiz 2013 smd.bwgrpS(18.1, 9.3, 30, 30, .05)
  # Wang 2019 smd.bwgrpS(24.86, 29.63, 37, 36, .238)
  mutate(
    rgv_sd_1 = ifelse(refid == 2985, 29.0, rgv_sd_1),
    rgv_sd_1 = ifelse(refid == 3154, 17.0, rgv_sd_1)
  ) %>%
  arrange(refid, arm_tx)

# use calc_mn_sd function (relies on meta) to calculate missing mean, sd, and log transformed
rgv_meta.dat <- calc_mn_sd(
  "arm_n",
  "rgv_mean_1",
  "rgv_sd_1",
  "rgv_median_1",
  "rgv_iqr_low_1",
  "rgv_iqr_up_1",
  "rgv_range_low_1",
  "rgv_range_up_1",
  "study",
  "arm_tx",
  refid = "refid",
  log_trans = TRUE,
  data = rgv_meta.dat) %>%
  bind_cols(., rgv_meta.dat %>% select(hr_gt6:hr_2))

# combine naguib 2001 CHO arms, plain and log
naguib <- c(combine_contin(75, 75, 27.08, 23.38, 20.35, 11.547), 
            combine_contin(75, 75, 3.074936, 3.042742, 0.6691179, 0.4672016)[2:3])
rgv_meta.dat[36, c(3:7)] <- naguib
rm(naguib)
# delete arm used to combine
rgv_meta.dat <- as_tibble(rgv_meta.dat) %>% 
  filter(!(arm_tx == "CHO/Simp" & n == 75))

## (updated 2021/05/15 07:56) add ingestion time ----------------------------
rgv_meta.dat <- rgv_meta.dat %>%
  mutate(
    arm_tx = fct_drop(arm_tx),
    hr_2_liq = ifelse(!is.na(hr_2), "-2", "-gt2"),
    temp_arm = as.character(arm_tx),
    arm_tx_time = ifelse(arm_tx %in% c("CHO/Comp", "CHO/Simp", "CHO/Comp/Prot", "CHO/Simp/Prot", "Clear", "Water"),
      paste0(temp_arm, hr_2_liq), temp_arm),
    arm_tx_time = factor(arm_tx_time,
      levels = c(
        "Fasting",
        "Water-2",
        "Clear-2",
        # "Placebo",
        "CHO/Simp-2",
        "CHO/Comp/Prot-2",
        # "CHO/Simp/Prot-2",
        "CHO/Comp-2",
        "CHO/Simp-gt2",
        "CHO/Comp/Prot-gt2",
        "CHO/Comp-gt2"
      )
    ),
    # arm_tx_sim_com = fct_collapse(arm_tx, "Protein" = c("CHO/Simp/Prot", "CHO/Comp/Prot")),
    arm_tx_cho_prot = fct_collapse(arm_tx, # 3 categories
      "CHO" = c("CHO/Simp", "CHO/Comp"),
      "Protein" = c("CHO/Comp/Prot")),
    arm_tx_time_no_prot = fct_collapse(arm_tx_time, # 3 categories
      # Fasting = "Fasting",
      # Water = "Water",
      # Placebo = "Placebo",
      # "CHO/Simp"   = c("CHO/Simp-2", "CHO/Simp-gt2"),
      # "CHO/Comp-2"   = c("CHO/Comp-2", "CHO/Comp/Prot-2"),
      # "CHO/Comp-gt2" = c("CHO/Comp-gt2", "CHO/Comp/Prot-gt2")
      "CHO-gt2"   = c("CHO/Comp/Prot-gt2", "CHO/Simp-gt2", "CHO/Comp-gt2"),
      "CHO-2"   = c("CHO/Simp-2", "CHO/Comp-2", "CHO/Comp/Prot-2")
    ),
    arm_tx_time2 = fct_collapse(arm_tx_time, # 2 categories
      # Fasting = "Fasting",
      # Water = "Water",
      # Placebo = "Placebo",
      # "CHO/Simp"   = c("CHO/Simp-2", "CHO/Simp-gt2", "CHO/Simp/Prot-2"),
      "CHO 2hr"   = c("CHO/Comp-2", "CHO/Simp-2"),
      "CHO >2hr" = c("CHO/Comp-gt2", "CHO/Simp-gt2"),
      "Protein" = c("CHO/Comp/Prot-gt2", "CHO/Comp/Prot-2")
    )
  ) %>% 
  select(-c(hr_2_liq, temp_arm))

## (updated 2021/05/15 07:25) for pairwise not run ------------------------
# rgv_meta.dat <- rgv_meta.dat %>%
#   mutate(arm_tx = ifelse(grepl("CHO", arm_tx), "CHO", "Fasting")) %>%
#   pivot_wider(
#     id_cols = c("study"),
#     names_from = c("arm_tx"),
#     values_from = c("n", "mean", "sd", "mean_log", "sd_log")
#   )
```


# **Residual Gastric Volume**

## *Network Characteristics*

<br/>

<font size = 4> Figure `r figure_n`. Network plot of comparators in included studies. </font>

```{r rgvNMA, include = TRUE, echo = FALSE, fig.align = 'left', fig.width = 6, fig.height = 5}
## (updated 2021/05/15 08:48) network characteristics ---------------------
# par(mfrow = c(1, 3))
# for (outcome in c("arm_tx", "arm_tx_cho_prot", "arm_tx_time2")){
# rgv.data <- data.prep(
#   arm.data = rgv_meta.dat,
#   varname.t = outcome,
#   varname.s = "study"
# )
# 
# net.plot(rgv.data,
#   node.scale = 3,
#   edge.scale = 1.5,
#   label.offset2 = 1,
#   label.offset1 = 0,
#   study.counts = TRUE,
#   node.lab.cex = 1.6,
#   edge.lab.cex = 1.8
# )
# }

#### (updated 2021/05/15 10:30) ----------------------------

## (updated 2021/03/16 07:13) rgvSimpComp BUGSnet -------------------------
# BUGSnet
rgv.data <- data.prep(
  arm.data = rgv_meta.dat,
  varname.t = "arm_tx",
  varname.s = "study"
)

net.plot(rgv.data,
  node.scale = 3,
  edge.scale = 1.5,
  label.offset2 = 1,
  label.offset1 = 4,
  study.counts = TRUE,
  node.lab.cex = 1,
  edge.lab.cex = 1
)
```

<font size = 4> Table `r table_n`. Network characteristics. </font>

```{r netTableRgv}
net_table <- net.tab(rgv.data,
  outcome = "mean",
  N = "n",
  type.outcome = "continuous"
)

net_table$network %>%
  kbl(booktabs = T, align =  c("lr"), escape = FALSE, col.names = c("Characteristic", "Value")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "30em") %>%
  column_spec(2, width = "6em")

table_n <- tab_inc()
```

<br/>

<font size = 4> Table `r table_n`. Number of studies and patients in comparisons represented in the network. </font>

```{r}
net_table$comparison %>%
  kbl(booktabs = T, align =  c("lcc"), escape = FALSE, col.names = c("Comparison", "Studies", "Patients")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "20em") %>%
  column_spec(c(2,3), width = "6em")

table_n <- tab_inc()
```

<br/>

<font size = 4> Table `r table_n`. Mean and ranges of residual gastric volumes across included studies. </font>

```{r}
net_table$intervention %>%
  mutate(across(min.outcome:av.outcome, ~ round(., 1))) %>% 
  kbl(booktabs = T, align =  c("lccccc"), escape = FALSE, col.names = c("Liquid", "Studies", "Patients", "RGV min", "RGV max", "Mean RGV")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "14em") %>%
  column_spec(c(2:5), width = "6em")

table_n <- tab_inc()
```

<br/>

## *Results Summary*

```{r rgvNMAmodel, include = FALSE, echo = FALSE, messages = FALSE}
rgv_re <- nma.model(
  data = rgv.data,
  outcome = "mean",
  N = "n",
  sd = "sd",
  reference = "Fasting",
  family = "normal",
  link = "identity",
  effects = "random",
  type = "consistency"
)

re_results <- nma.run(rgv_re,
  n.adapt = 101,
  n.burnin = 20000,
  n.iter = 50000
)

## (updated 2021/05/15 10:38) not run sucra and league --------------------
# nma.diag(re_results)
# temp <- nma.diag(re_results)
# temp$gelman.rubin
#       Point est. Upper C.I.
# d[2]   1.0001634  1.0006036
# d[3]   1.0000197  1.0000737
# d[4]   0.9999903  0.9999998
# d[5]   1.0000211  1.0000850
# d[6]   1.0000268  1.0001322
# sigma  0.9999996  1.0000330
# rm(temp)
# SUCRA Plot
# (sucra.out <- nma.rank(re_results, largerbetter = FALSE, sucra.palette = "Set1"))
# League table
df <- nma.league(re_results,
  central.tdcy = "median",
  # order = sucra.out$order,
  log.scale = FALSE,
  low.colour = "springgreen4",
  mid.colour = "white",
  high.colour = "red"
)

```

<br/>

<font size = 4> Table `r table_n`. Estimated mean differences in residual gastric volume comapred with complex carbohydrate containing liquids. </font>

```{r rgvKable, include = TRUE}
## (updated 2021/05/15 10:44) table of estimates with cho_complex ---------
order_dat <- tibble(Treatment = c("CHO/Simp", "CHO/Comp/Prot", "Clear", "Water", "Fasting"))

temp <- df$heatplot$data %>%
  filter(Comparator == "CHO/Comp") %>%
  as_tibble()

rgv_table_kbl <- left_join(order_dat, temp, by = "Treatment") %>%
  mutate(across(c(ct.stat:uci), ~ format(round(.x, 1), nsmall = 1)),
    cri = paste0("(", lci, " to ", uci, ")"),
    # md_cri = paste0(ct.stat, " (", lci, " to ", uci, ")"),
    cri = str_replace(cri, "\\( ", "\\("),
    compared = paste0(Treatment, " vs. ", Comparator)
  ) %>%
  select(compared, ct.stat, cri)

rgv_table_kbl %>%
  kbl(
    booktabs = T, align = c("lcl"), escape = FALSE,
    col.names = c("Comparison", "MD (mL)", "  (95% CrI)")
  ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(bootstrap_options = opt_boot, full_width = FALSE, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  column_spec(1, width = "14em") %>%
  column_spec(2, width = "6em") %>%
  column_spec(3, width = "7em") %>%
  footnote(
    general = "MD: mean difference; CrI: credible interval; CHO: carbohydrate; Comp: complex; Prot: Protein",
    # alphabet = c(),
    general_title = "",
    footnote_as_chunk = FALSE
  )

table_n <- tab_inc()
```

<font size = 2> A sensitivity analysis with residual gastric volumes transformed to the log scale to account for skewed distributions yielded consistent results.  </font>

<br/>

<font size = 4> Figure `r figure_n`. Forest plot of estimated mean differences in residual gastric volume comapred with complex carbohydrate containing liquids. </font>

```{r rgvForest, include = TRUE, fig.align = 'left', fig.width = 7, fig.height = 4}
## (updated 2021/05/15 10:44) forest plot ---------------------------------
nma.forest(re_results, comparator = "CHO/Comp") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12))

```

# **Residual Gastric Volume Time** 
```{r rgvTimeNMAmodel, include = TRUE, echo = FALSE, messages = FALSE}
## (updated 2021/05/15 13:47) TimeNMAmodel --------------------------------
rgv_meta_time.dat <- rgv_meta.dat %>% 
  filter(!study %in% c("Braga 2012","Karimian 2020"))

rgv_time.data <- data.prep(
  arm.data = rgv_meta_time.dat,
  varname.t = "arm_tx_time_no_prot",
  varname.s = "study"
)

net.plot(rgv_time.data,
  node.scale = 3,
  edge.scale = 1.5,
  label.offset2 = 1,
  label.offset1 = 4,
  study.counts = TRUE,
  node.lab.cex = 1,
  edge.lab.cex = 1
)

rgv_re <- nma.model(
  data = rgv_time.data,
  outcome = "mean",
  N = "n",
  sd = "sd",
  reference = "Fasting",
  family = "normal",
  link = "identity",
  effects = "random",
  type = "consistency"
)

re_results <- nma.run(rgv_re,
  n.adapt = 101,
  n.burnin = 20000,
  n.iter = 50000
)

df <- nma.league(re_results,
  central.tdcy = "median",
  # order = sucra.out$order,
  log.scale = FALSE,
  low.colour = "springgreen4",
  mid.colour = "white",
  high.colour = "red"
)

```

<br/>

<font size = 4> Table `r table_n`. Estimated mean differences in residual gastric volume compared with complex carbohydrate containing liquids. </font>

```{r rgvTimeKable, include = TRUE}
## (updated 2021/05/15 10:44) table of estimates with cho_complex ---------
order_dat <- tibble(Treatment = c("CHO-gt2", "Clear-2", "Water-2", "Fasting"))

temp <- df$heatplot$data %>%
  filter(Comparator == "CHO-2") %>%
  as_tibble()

rgv_table_kbl <- left_join(order_dat, temp, by = "Treatment") %>%
  mutate(across(c(ct.stat:uci), ~ format(round(.x, 1), nsmall = 1)),
    cri = paste0("(", lci, " to ", uci, ")"),
    # md_cri = paste0(ct.stat, " (", lci, " to ", uci, ")"),
    cri = str_replace(cri, "\\( ", "\\("),
    compared = paste0(Treatment, " vs. ", Comparator)
  ) %>%
  select(compared, ct.stat, cri)

rgv_table_kbl %>%
  kbl(
    booktabs = T, align = c("lcl"), escape = FALSE,
    col.names = c("Comparison", "MD (mL)", "  (95% CrI)")
  ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(bootstrap_options = opt_boot, full_width = FALSE, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  column_spec(1, width = "14em") %>%
  column_spec(2, width = "6em") %>%
  column_spec(3, width = "7em") %>%
  footnote(
    general = "MD: mean difference; CrI: credible interval; CHO: carbohydrate; Comp: complex; Prot: Protein",
    # alphabet = c(),
    general_title = "",
    footnote_as_chunk = FALSE
  )

table_n <- tab_inc()
```

<br/>

<font size = 4> Figure `r figure_n`. Forest plot of estimated mean differences in residual gastric volume comapred with complex carbohydrate containing liquids. </font>

```{r rgvTimeForest, include = TRUE, fig.align = 'left', fig.width = 7, fig.height = 4}
## (updated 2021/05/15 10:44) forest plot ---------------------------------
nma.forest(re_results, comparator = "CHO-2") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12))

```


## *Network Characteristics*

<br/>

<font size = 4> Table `r table_n`. Network characteristics. </font>

```{r netTableTimeRgv}
net_table <- net.tab(rgv_time.data,
  outcome = "mean",
  N = "n",
  type.outcome = "continuous"
)

net_table$network %>%
  kbl(booktabs = T, align =  c("lr"), escape = FALSE, col.names = c("Characteristic", "Value")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "30em") %>%
  column_spec(2, width = "6em")

table_n <- tab_inc()
```

<br/>

<font size = 4> Table `r table_n`. Number of studies and patients in comparisons represented in the network. </font>

```{r}
net_table$comparison %>%
  kbl(booktabs = T, align =  c("lcc"), escape = FALSE, col.names = c("Comparison", "Studies", "Patients")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "20em") %>%
  column_spec(c(2,3), width = "6em")

table_n <- tab_inc()
```

<br/>

<font size = 4> Table `r table_n`. Mean and ranges of residual gastric volumes across included studies. </font>

```{r}
net_table$intervention %>%
  mutate(across(min.outcome:av.outcome, ~ round(., 1))) %>% 
  kbl(booktabs = T, align =  c("lccccc"), escape = FALSE, col.names = c("Liquid", "Studies", "Patients", "RGV min", "RGV max", "Mean RGV")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "14em") %>%
  column_spec(c(2:5), width = "6em")

table_n <- tab_inc()
```

<br/>


# Liquid Detail 

## Carbohydrate Amounts

<br/>

<font size = 4> Figure `r figure_n`. Total carbohydrate consumed prior to surgery according to type and combined with protein in adult trials. </font>

```{r choAmtsFigure, include = TRUE, echo = FALSE, include = TRUE, fig.width = 6, fig.height = 2.5, fig.align = "left"}
## * (updated 2021/03/10 13:29) cho_amts adult rct ----------------------------
cho_amts <- study_arm_cho.dat %>%
  filter(design == "rct", age == "Adult", surg_nosurg == "surgical") %>%
  filter(arm_tx %in% c("CHO/Comp/Prot", "CHO/Simp/Prot", "CHO/Comp", "CHO/Simp")) %>%
  filter(refid %in% cho_refids) %>% 
  mutate(
    amtingest2 = replace_na(amtingest2, 0),
    amtingest1 = replace_na(amtingest1, 0),
    tot_vol = amtingest2 + amtingest1,
    prot_gm_all = prot_conc * tot_vol / 100,
    cho_gm_all = cho_conc * tot_vol / 100,
    cho_gm_1 = cho_conc * amtingest1 / 100,
    cho_gm_2 = cho_conc * amtingest2 / 100,
    cho_gm_all = ifelse(is.na(cho_gm_all), cho_gm, cho_gm_all),
    prot_gm_all = ifelse(is.na(prot_gm_all), protein_gm, prot_gm_all)) %>% 
  select(refid, study, arm_tx, arm, cho_gm, protein_gm, prot_gm_all, cho_gm_all, cho_gm_1, cho_gm_2, amtingest1, amtingest2, prot_conc, cho_conc)

## figure of cho amounts ####
theme <- theme_set(theme_minimal())
theme <- theme_update(
  legend.position = "none",
  axis.text = element_text(size = 14), 
  axis.title = element_text(size = 14),
  axis.line.y = element_line(size = 0, colour = "white", linetype = "solid"), 
  axis.ticks = element_line(colour = "gray72"),
  panel.grid.major = element_line(colour = "gray80"))
par(mar = c(6, 5, 2, 2), oma = c(1, 1, 2, 2))

cho_amts %>% ggplot(aes(cho_gm_all, arm_tx)) +
  geom_boxplot(varwidth = TRUE) +
  geom_jitter(width = .2, alpha = 0.8) +
  scale_y_discrete(name = "") +
  scale_x_continuous(name = "Grams", limits = c(0, 200)) 
  # + theme(axis.text=element_text(size = 16),
  #       axis.title=element_text(size = 16,face="bold"))

figure_n <- fig_inc()
```
<font size = 3> CHO: carbohydrate; Comp: complex; Simp: simple; Prot: protein. </font>

<br/>
