---
title: "Network Meta-Analyses — Carbohydrates"
date: "`r format(Sys.time(), '%d %B, %Y (%H:%M)')`"
output:
  html_document:
    font: Source Sans Pro
    theme: cosmo
    anchor_sections: TRUE
    toc_float: 
      collapsed: FALSE
      smooth_scroll: TRUE
    toc_depth: 3
    linkcolor: blue
  latex_engine: xelatex
  pdf_document:
    toc: yes
    toc_depth: '3'
bibliography: cho_nma.bib
csl: jama.csl
link-citations: yes
workflowr:
  suppress_report: false
nocite: '@*'
---

#### 

```{=html}
<style type="text/css">
  body{
  font-size: 9.5pt;
  font-family: Source Sans Pro;
}

.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
  color: #090909;
  background-color: #0909093d;
}

.lightable-classic.lightable-hover tbody tr:hover {
    background-color: #c5d4e4;
}
.btn-xs, .btn-group-xs>.btn { 
display: none; 
}

.btn-default {
display: none !important; 
}
</style>
```

```{r setupData, include = FALSE}
## * (updated 2021/01/06 11:34) preliminaries -------------------------------
library(multinma)
library(BUGSnet)
# functions
now <- function() {str_replace_all(format(Sys.time(), "%m %d %y_%H-%M"), " ", "_")}
m_table <- function(data, x){
  group_by(data, arm_tx) %>%
    summarise(tx_count = n())
}
# replace_mg <- function(a, b) {if_else(is.na(a) & !is.na(b), b, a)}
tab_inc <- function() {
  table_n <- table_n + 1
  table_n}
fig_inc <- function() {
  figure_n <- figure_n + 1
  figure_n}
table_n <- 1
figure_n <- 1

## * get current files; select cho refids; directory must have unique ####
source("code/readFiles_120220_wfr.R")

# study char cho refs 
study_char_cho.dat <- study_char.dat %>% 
  filter(refid %in% cho_refids)

# study arm protein refs 
study_arm_cho.dat <- study_arm.dat %>%
  filter(refid %in% cho_refids) %>%
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>%
  # remove duplicate arms tewari 2019 (refid 716)
  filter(!(refid == 716 & arm_n == 8)) %>%
  mutate(
    study = ifelse(refid == 2564 & grepl("colorectal", notes_arm), paste(study, "(CR)"), study),
    study = ifelse(refid == 2564 & grepl("cholecystectomy", notes_arm), paste(study, "(CH)"), study),
    study = ifelse(refid == 1926 & grepl("general", notes_arm), paste(study, "(gen)"), study),
    study = ifelse(refid == 1926 & grepl("epidural", notes_arm), paste(study, "(epi)"), study)
  ) %>%
  mutate(tx = factor(arm_tx,
    levels = c(
      "fasting",
      "water",
      "placebo",
      "other_clear",
      "milk",
      "prot",
      "prot_simp",
      "prot_comp",
      "cho_simp",
      "cho_comp"
    )
  )) %>%
  mutate(
    tx = fct_recode(tx,
      Fasting = "fasting",
      Water = "water",
      Placebo = "placebo",
      Clear = "other_clear",
      Milk = "milk",
      "Prot" = "prot",
      "CHO/Comp/Prot" = "prot_comp",
      "CHO/Simp/Prot" = "prot_simp",
      "CHO/Simp" = "cho_simp",
      "CHO/Comp" = "cho_comp",
    ),
    arm_tx = tx
  ) %>%
  select(-tx) 

# tewari 2019
study_arm_cho_tewari.dat <- study_arm.dat %>% 
  filter(refid == 716)

## * dichot factor treatment arm outcome type for ordering (updated 2020/12/12 12:04) ####
# save dichot.dat 
dichot_cho.dat <- dichot.dat %>% 
  filter(refid %in% cho_refids) %>% 
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>% 
  # factor arm_tx levels for cho studies
  mutate(tx = factor(arm_tx,
  levels = c(
    "fasting",
    "water",
    "placebo",
    "other_clear",
    "milk",
    "prot",
    "prot_simp",
    "prot_comp",
    "cho_simp",
    "cho_comp"))) %>%
  mutate(tx = fct_recode(tx,
    Fasting = "fasting",
    Water = "water",
    Placebo = "placebo",
    Clear = "other_clear",
    Milk = "milk",
    "Prot" = "prot",
    "CHO/Simp/Prot" = "prot_simp",
    "CHO/Comp/Prot" = "prot_comp",
    "CHO/Simp" = "cho_simp",
    "CHO/Comp" = "cho_comp",),
    arm_tx = tx) %>%
  select(-tx)

dichot_cho.dat %>% tabyl(arm_tx)

## * contin factor treatment arm outcome type for ordering (updated 2020/12/12 12:04) ####
contin_cho.dat <- contin.dat %>% 
  filter(refid %in% cho_refids) %>% 
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>%
  # factor arm_tx levels for cho studies
  mutate(tx = factor(arm_tx,
  levels = c(
    "fasting",
    "water",
    "placebo",
    "other_clear",
    "milk",
    "prot",
    "prot_simp",
    "prot_comp",
    "cho_simp",
    "cho_comp"))) %>%
  mutate(tx = fct_recode(tx,
    Fasting = "fasting",
    Water = "water",
    Placebo = "placebo",
    Clear = "other_clear",
    Milk = "milk",
    "Prot" = "prot",
    "CHO/Simp/Prot" = "prot_simp",
    "CHO/Comp/Prot" = "prot_comp",
    "CHO/Simp" = "cho_simp",
    "CHO/Comp" = "cho_comp",),
    arm_tx = tx) %>%
  select(-tx)

contin_cho.dat %>% tabyl(arm_tx)

## * likert factor treatment arm outcome type for ordering (updated 2020/12/12 12:04) ####
likert_cho.dat <- likert.dat %>% 
  filter(refid %in% cho_refids) %>% 
  # remove Awad other, no "other" category
  filter(arm_tx != "other") %>%
  # factor arm_tx levels for cho studies
  mutate(tx = factor(arm_tx,
  levels = c(
    "fasting",
    "water",
    "placebo",
    "other_clear",
    "milk",
    "prot",
    "prot_simp",
    "prot_comp",
    "cho_simp",
    "cho_comp"))) %>%
  mutate(tx = fct_recode(tx,
    Fasting = "fasting",
    Water = "water",
    Placebo = "placebo",
    Clear = "other_clear",
    Milk = "milk",
    "Prot" = "prot",
    "CHO/Simp/Prot" = "prot_simp",
    "CHO/Comp/Prot" = "prot_comp",
    "CHO/Simp" = "cho_simp",
    "CHO/Comp" = "cho_comp",),
    arm_tx = tx) %>%
  select(-tx)

likert_cho.dat %>% tabyl(arm_tx)

## * divide Sada 2014 and Celiksular 2016 into 2 studies ------------------
likert_cho.dat <- likert_cho.dat %>%
  mutate(
    study = ifelse(study == "Sada 2014" & grepl("cholecystectomy", out_l_notes), "Sada 2014 (CH)", study),
    study = ifelse(study == "Sada 2014" & !grepl("cholecystectomy", out_l_notes), "Sada 2014 (CR)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("general", out_l_notes), "Celiksular 2016 (G)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("epidural", out_l_notes), "Celiksular 2016 (E)", study)
  )

contin_cho.dat <- contin_cho.dat %>%
  mutate(
    study = ifelse(study == "Sada 2014" & grepl("cholecystectomy", notes_continuous), "Sada 2014 (CH)", study),
    study = ifelse(study == "Sada 2014" & !grepl("cholecystectomy", notes_continuous), "Sada 2014 (CR)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("general", notes_continuous), "Celiksular 2016 (G)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("epidural", notes_continuous), "Celiksular 2016 (E)", study)
  )

dichot_cho.dat <- dichot_cho.dat %>%
  mutate(
    study = ifelse(study == "Sada 2014" & grepl("cholecystectomy", out_dichot_notes), "Sada 2014 (CH)", study),
    study = ifelse(study == "Sada 2014" & !grepl("cholecystectomy", out_dichot_notes), "Sada 2014 (CR)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("general", out_dichot_notes), "Celiksular 2016 (G)", study),
    study = ifelse(study == "Celiksular 2016" & grepl("epidural", out_dichot_notes), "Celiksular 2016 (E)", study)
  )


## (updated 2021/07/05 10:39) rgv_meta_nma_studies ------------------------
rgv_meta_nma_studies <- contin_cho.dat %>%
  arrange(design, year, study) %>%
  filter(refid %in% cho_refids, surg_nosurg == "surgical") %>%
  filter(!is.na(rgv_us) | !is.na(rgv_imag) | !is.na(rgv_asp)) %>%
  filter(design == "rct", age == "Adult") %>%
  select(refid, study, year, design, arm_n, arm_tx, rgv_mean_1, rgv_sd_1, rgv_median_1:rgv_iqr_up_1, hr_gt6:hr_2) %>%
  select(refid, study) %>%
  distinct()
```

```{r rgvNmaData, include = TRUE} 
## (updated 2021/07/02 11:33) prepare rgv_meta.dat for nma ----------------
# arm_tx levels "Fasting", "Water", "Placebo", "Clear", "Milk", "Prot", "CHO/Simp/Prot", "CHO/Comp/Prot", "CHO/Simp", "CHO/Comp"
## (updated 2021/05/15 07:21) rgv_meta.dat --------------------------------
rgv_meta.dat <- contin_cho.dat %>%
  arrange(design, year, study) %>%
  filter(refid %in% cho_refids, surg_nosurg == "surgical") %>%
  filter(!is.na(rgv_us) | !is.na(rgv_imag) | !is.na(rgv_asp)) %>%
  filter(design == "rct", age == "Adult") %>%
  select(refid, study, year, design, arm_n, arm_tx, rgv_mean_1, rgv_sd_1, rgv_median_1:rgv_iqr_up_1, hr_gt6:hr_2) %>%
  # calculate sd not reported using p-values 
  mutate(
    rgv_sd_1 = ifelse(study == "Yilmaz 2013", sd_bwgrp(21.85, 19.15, 20, 20, .77), rgv_sd_1),
    rgv_sd_1 = ifelse(study == "Yildiz 2013", sd_bwgrp(18.1, 9.3, 30, 30, .05), rgv_sd_1)
  ) %>%
  arrange(refid, arm_tx)

# use calc_mn_sd function (relies on meta) to calculate if missing mean, sd, and log transformed
rgv_meta.dat <- calc_mn_sd(
  "arm_n",
  "rgv_mean_1",
  "rgv_sd_1",
  "rgv_median_1",
  "rgv_iqr_low_1",
  "rgv_iqr_up_1",
  "rgv_range_low_1",
  "rgv_range_up_1",
  "study",
  "arm_tx",
  refid = "refid",
  log_trans = TRUE,
  data = rgv_meta.dat) %>%
  bind_cols(., rgv_meta.dat %>% select(hr_gt6:hr_2))

## (updated 2021/05/15 07:56) add ingestion time ----------------------------
rgv_meta.dat <- rgv_meta.dat %>%
  mutate(
    arm_tx = fct_drop(arm_tx),
    hr_2_liq = ifelse(!is.na(hr_2), "-2", "-gt2"),
    temp_arm = as.character(arm_tx),
    arm_tx_time = ifelse(arm_tx %in% c("CHO/Comp", "CHO/Simp", "CHO/Comp/Prot", "CHO/Simp/Prot", "Clear", "Water"),
      paste0(temp_arm, hr_2_liq), temp_arm),
    arm_tx_time = factor(arm_tx_time,
      levels = c(
        "Fasting",
        "Water-2",
        "Clear-2",
        # "Placebo",
        "CHO/Simp-2",
        "CHO/Comp/Prot-2",
        # "CHO/Simp/Prot-2",
        "CHO/Comp-2",
        "CHO/Simp-gt2",
        "CHO/Comp/Prot-gt2",
        "CHO/Comp-gt2"
      )
    ),
    # arm_tx_sim_com = fct_collapse(arm_tx, "Protein" = c("CHO/Simp/Prot", "CHO/Comp/Prot")),
    arm_tx_cho_prot = fct_collapse(arm_tx, # 3 categories
      "CHO" = c("CHO/Simp", "CHO/Comp"),
      "Protein" = c("CHO/Comp/Prot")),
    arm_tx_time_no_prot = fct_collapse(arm_tx_time, # 3 categories
      # Fasting = "Fasting",
      # Water = "Water",
      # Placebo = "Placebo",
      # "CHO/Simp"   = c("CHO/Simp-2", "CHO/Simp-gt2"),
      # "CHO/Comp-2"   = c("CHO/Comp-2", "CHO/Comp/Prot-2"),
      # "CHO/Comp-gt2" = c("CHO/Comp-gt2", "CHO/Comp/Prot-gt2")
      "CHO-gt2"   = c("CHO/Comp/Prot-gt2", "CHO/Simp-gt2", "CHO/Comp-gt2"),
      "CHO-2"   = c("CHO/Simp-2", "CHO/Comp-2", "CHO/Comp/Prot-2")
    ),
    arm_tx_time2 = fct_collapse(arm_tx_time, # 2 categories
      # Fasting = "Fasting",
      # Water = "Water",
      # Placebo = "Placebo",
      # "CHO/Simp"   = c("CHO/Simp-2", "CHO/Simp-gt2", "CHO/Simp/Prot-2"),
      "CHO 2hr"   = c("CHO/Comp-2", "CHO/Simp-2"),
      "CHO >2hr" = c("CHO/Comp-gt2", "CHO/Simp-gt2"),
      "Protein" = c("CHO/Comp/Prot-gt2", "CHO/Comp/Prot-2")
    )
  ) %>% 
  select(-c(hr_2_liq, temp_arm))

## not run (2021/05/15 07:25) for pairwise --------------------------------
#rgv_meta.dat <- rgv_meta.dat %>%
#  mutate(arm_tx = ifelse(grepl("CHO", arm_tx), "CHO", "Fasting")) %>%
#  pivot_wider(
#    id_cols = c("study"),
#    names_from = c("arm_tx"),
#    values_from = c("n", "mean", "sd", "mean_log", "sd_log")
#  )
```

# **Included Studies**

<br/>

<font size = 4> Table `r table_n`. Randomized controlled trials included in the network meta-analysis (all adults undergoing surgery).  See [References](#references) for citations. </font>

```{r studiesIncludedList, echo = FALSE, include = TRUE}
## * (updated 2021/03/01 09:25) included_studies ----------------------------
included_studies <- study_char_cho.dat %>%
  filter(refid %in% rgv_meta_nma_studies$refid) %>% 
  select(refid, study, year, age, surg_nosurg, design, n_analyze, country, centers, non_vh_hdi, surg_type) %>%
  mutate(
    surg = case_when(
      str_detect(surg_type, "[Gg]yn|[Hh]ysterectomy") ~ "Gyn",
      str_detect(surg_type, "craniotomy") ~ "Neurosurgical",
      str_detect(surg_type, "maxillofacial") ~ "Maxillofacial",
      str_detect(surg_type, "[Hh]ernia|herniorrhaphy") ~ "Herniorrhaphy",
      str_detect(surg_type, "prostatectomy") ~ "Prostatectomy",
      str_detect(surg_type, "ariety") ~ "Various",
      str_detect(surg_type, "various|Hirschsprung's") ~ "Various",
      str_detect(surg_type, "CABG|[Cc]oronary artery bypass surgery") ~ "CABG",
      str_detect(surg_type, "Cardiac surgery") ~ "Cardiac",
      str_detect(surg_type, "[Cc]olorectal|[Bb]owel|colon resection") ~ "Colorectal",
      str_detect(surg_type, "[Tt]hyroid|hip joint") ~ "Thyroid",
      str_detect(surg_type, "[Oo]rthopedic|discectomy") ~ "Orthopedic",
      str_detect(surg_type, "total hip or total knee replacement|[Tt]otal hip|[Tt]otal knee|Hip joint|hip replacement") ~ "TKA/THA",
      str_detect(surg_type, "cholecystectomy") ~ "Cholecystectomy",
      str_detect(surg_type, "pancreaticoduodenectomy") ~ "Whipple",
      str_detect(surg_type, "[Gg]ast|Bowel") ~ "Other GI",
      str_detect(surg_type, "not appli") ~ "None (healthy)",
      str_detect(surg_type, "oral") ~ "Oral",
      str_detect(surg_type, "minor") ~ "Misc minor",
      str_detect(surg_type, "None") ~ "None (healthy)",
      str_detect(surg_type, "abdomin") ~ "Other GI",
      str_detect(surg_type, "[Cc]esarean") ~ "Cesarean delivery",
      str_detect(surg_type, "[Cc]aesarian") ~ "Cesarean delivery",
      str_detect(surg_type, "[Ll]abor") ~ "Labor",
      str_detect(surg_type, "thoracotomy") ~ "Thoracic",
      str_detect(surg_type, "healthy") ~ "None (healthy)",
      str_detect(surg_type, "Elective") ~ "Elective",
      str_detect(surg_type, "adenotonsillectomy") ~ "Adenotonsillectomy",
      str_detect(surg_type, "[Bb]reast") ~ "Breast",
      str_detect(surg_type, "Endso|endoscopic|colonoscopy") ~ "Endoscopic"
    ),
    surg_nosurg = ifelse(surg_nosurg == "non-surgical", "None", surg_nosurg),
    study = ifelse(grepl("Marquini", study), str_c(study, "^b^"), study),
    country = ifelse(grepl("USA", country), "USA", country),
    country = ifelse(grepl("UK", country), "UK", country),
    country = ifelse(!is.na(non_vh_hdi), str_c(country, "^a^"), country),
    n_analyze = ifelse(refid == 2009, 9, n_analyze)
  ) 

## (updated 2021/03/01 09:24) adult surgical ------------------------------
included_studies %>% 
  arrange(age, design, surg, country, year, study) %>%
  filter(age == "Adult", surg_nosurg == "surgical") %>% 
  select(refid, study, n_analyze, centers, country, surg) %>%
  kbl(
    booktabs = T, align = c("llccll"), # format = "latex",
    col.names = c("        ID", "Study", "Analyzed (N)", "Centers", "Country^a^", "Surgery")
  ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(bootstrap_options = opt_boot, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  column_spec(1, width = "6em") %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "6em") %>%
  column_spec(4, width = "6em") %>%
  column_spec(5, width = "10em") %>%
  column_spec(6, width = "12em") %>%
  pack_top(., "RCT", 1, 15) %>%
  footnote(
    general = "RCT: randomized controlled trial.",
    alphabet = c(
      "Non very high [Human Development Index](https://en.wikipedia.org/wiki/List_of_countries_by_Human_Development_Index) country."),
    general_title = "",
    footnote_as_chunk = FALSE
  )

table_n <- tab_inc()
```

<br/>


# **Residual Gastric Volume**

## *Network Characteristics*

<br/>

<font size = 4> Figure `r figure_n`. Network plot of comparators in included studies. </font>

```{r rgvNMA, include = TRUE, echo = FALSE, fig.align = 'left', fig.width = 5, fig.height = 5}
## not run networks including time and protein ----------------------------
# par(mfrow = c(1, 3))
# for (outcome in c("arm_tx", "arm_tx_cho_prot", "arm_tx_time2")){
# rgv.data <- data.prep(
#   arm.data = rgv_meta.dat,
#   varname.t = outcome,
#   varname.s = "study"
# )
# 
# net.plot(rgv.data,
#   node.scale = 3,
#   edge.scale = 1.5,
#   label.offset2 = 1,
#   label.offset1 = 0,
#   study.counts = TRUE,
#   node.lab.cex = 1.6,
#   edge.lab.cex = 1.8
# )
# }

## (updated 2021/03/16 07:13) rgvSimpComp BUGSnet -------------------------
# for log
rgv_meta_log.dat <- rgv_meta.dat %>% 
  select(-mean, -sd) %>% 
  rename(mean = mean_log, sd = sd_log)

rgv.data <- data.prep(
  arm.data = rgv_meta.dat,
  # arm.data = rgv_meta_log.dat,
  varname.t = "arm_tx",
  varname.s = "study"
)

net.plot(rgv.data,
  node.scale = 3,
  edge.scale = 1.5,
  # label.offset2 = 3,
  # label.offset1 = 6,
  study.counts = TRUE,
  node.lab.cex = 1.3,
  edge.lab.cex = 1.2
)

# net.plot(rgv.data,
#   node.scale = 3,
#   edge.scale = 1.5,
#   # label.offset2 = 3,
#   # label.offset1 = 6,
#   # study.counts = TRUE,
#   node.lab.cex = 1.3,
#   edge.lab.cex = .5,
#   flag = "CHO/Comp/Prot"
# )
figure_n <- fig_inc()

```

<font size = 4> Table `r table_n`. Network characteristics. </font>

```{r netTableRgv}
net_table <- net.tab(rgv.data,
  outcome = "mean",
  N = "n",
  type.outcome = "continuous"
)

net_table$network %>%
  mutate(Characteristic = ifelse(Characteristic == "Average Outcome", "Average Residual Gastric Volume (mL)", Characteristic),
         Value = ifelse(Value == 21.89, 21.9, Value)) %>% 
  kbl(booktabs = T, align =  c("lr"), escape = FALSE, col.names = c("Characteristic", "Value")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "30em") %>%
  column_spec(2, width = "6em")

table_n <- tab_inc()
```

<br/>

<font size = 4> Table `r table_n`. Number of studies and patients in comparisons represented in the network. </font>

```{r}
net_table$comparison %>%
  kbl(booktabs = T, align =  c("lcc"), escape = FALSE, col.names = c("Comparison", "Studies", "Patients")) %>% 
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "20em") %>%
  column_spec(c(2,3), width = "6em") %>% 
  footnote(
    general = "CHO: carbohydrate; Comp: complex; Prot: protein.",
    # alphabet = c("Footnote a", "Footnote b"),
    general_title = "",
    footnote_as_chunk = FALSE
  )

table_n <- tab_inc()
```

<br/>

<font size = 4> Table `r table_n`. Mean and ranges of residual gastric volumes across included studies. </font>

```{r}
net_table$intervention %>%
  mutate(across(min.outcome:av.outcome, ~ round(., 1))) %>% 
  kbl(booktabs = T, align =  c("lccccc"), escape = FALSE, col.names = c("Liquid", "Studies", "Patients", "min", "max", "mean")) %>% 
  add_header_above(c(" " = 3, "Residual Gastric Volume (mL)" = 3), line = TRUE, bold = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  kable_styling(full_width = FALSE, bootstrap_options = opt_boot, position = "left") %>%
  column_spec(1, width = "14em") %>%
  column_spec(c(2:5), width = "6em") %>% 
    footnote(
    general = "CHO: carbohydrate; Comp: complex; Prot: protein; RGV: residual gastric volume; min: minimum; max: maximum.",
    # alphabet = c("Footnote a", "Footnote b"),
    general_title = "",
    footnote_as_chunk = FALSE
  )

table_n <- tab_inc()
```

<br/>

## *Results Summary*

```{r rgvNMAmodel, include = FALSE, echo = FALSE, messages = FALSE}
rgv_re <- nma.model(
  data = rgv.data,
  outcome = "mean",
  N = "n",
  sd = "sd",
  reference = "Fasting",
  family = "normal",
  link = "identity",
  effects = "random",
  type = "consistency",
  # prior.sigma = "dnorm(0,1) T(0,)"
)

set.seed(12345)
re_results <- nma.run(rgv_re,
  n.adapt = 101,
  n.burnin = 20000,
  n.iter = 50000
)

## (updated 2021/05/15 10:38) not run sucra and league --------------------
# # nma.diag(re_results)
# $gelman.rubin
# $psrf
#       Point est. Upper C.I.
# d[2]    1.000252   1.000931
# d[3]    1.000356   1.001321
# d[4]    1.000011   1.000023
# d[5]    1.000088   1.000311
# sigma   1.000130   1.000317
# 
# $mpsrf
# [1] 1.000376

# nma.fit(re_results)

# sigma        2.930e+00  4.202295  5.072117  6.141869  9.013046
# tau 95% CrI 5.1 (2.9, 9.0)

# SUCRA Plot
# (sucra.out <- nma.rank(re_results, largerbetter = FALSE, sucra.palette = "Set1"))
```

<br/>

<font size = 4> Table `r table_n`. Mean differences in residual gastric volumes (mL) and 95% CrI from network meta-analysis. Comparisons are top to bottom below the diagonal (eg, Fasting vs. CHO/Comp is 0.71 mL) and bottom to top above the diagonal (eg, CHO/Comp vs. Fasting is -0.71 mL). Note that differences were not identified betwen any comparators, nor were the 95% CrI bounds consistent with potential clinically important differences. </font>


```{r leagueTable, include = TRUE}
# league table
df <- nma.league(re_results,
  central.tdcy = "median",
  # order = sucra.out$order,
  log.scale = FALSE,
  low.colour = "white",
  mid.colour = "white",
  high.colour = "white"
)

as_tibble(df$table) %>%
  mutate(
    across(Fasting:Water, ~ str_replace(., " \\(", "<br/>\\(")),
    across(Fasting:Water, ~ if_else(.x %in% c("Fasting", "CHO/Comp", "CHO/Comp/Prot", "CHO/Simp", "Water"),
      cell_spec(.x, bold = TRUE), .x
    ))
  ) %>%
  kbl(
    booktabs = T, align = c("ccccc"), escape = FALSE,
    col.names = c(rep(" ", 5))
  ) %>%
  kable_styling(bootstrap_options = c("hover"), full_width = FALSE, position = "left") %>%
  row_spec(1:5, extra_css = "vertical-align: middle !important; border-top: 1px solid LightGray; border-bottom: 1px solid LightGray;") %>% 
  footnote(
    general = "𝞃 = 5.1 mL (95% CrI: 2.9, 9.0)",
    # alphabet = c("Footnote a", "Footnote b", ...),
    general_title = "",
    footnote_as_chunk = FALSE
  )
table_n <- tab_inc()
  
```

<br/>

<font size = 4> Table `r table_n`. Estimated mean differences and 95% CrI in residual gastric volume compared with complex carbohydrate containing liquids. </font>

```{r rgvKable, include = TRUE}
## (updated 2021/05/15 10:44) table of estimates with cho_complex ---------
order_dat <- tibble(Treatment = c("CHO/Simp", "CHO/Comp/Prot", "Water", "Fasting"))

temp <- df$heatplot$data %>%
  filter(Comparator == "CHO/Comp") %>%
  as_tibble()

rgv_table_kbl <- left_join(order_dat, temp, by = "Treatment") %>%
  mutate(across(c(ct.stat:uci), ~ format(round(.x, 1), nsmall = 2)),
    cri = paste0("(", lci, " to ", uci, ")"),
    # md_cri = paste0(ct.stat, " (", lci, " to ", uci, ")"),
    cri = str_replace(cri, "\\( ", "\\("),
    compared = paste0(Treatment, " vs. ", Comparator)
  ) %>%
  select(compared, ct.stat, cri)

rgv_table_kbl %>%
  kbl(
    booktabs = T, align = c("lcl"), escape = FALSE,
    col.names = c("Comparison", "MD (mL)", "  (95% CrI)")
  ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(bootstrap_options = opt_boot, full_width = FALSE, position = "left") %>%
  kable_classic(full_width = FALSE, html_font = opt_font, position = "left", "hover") %>%
  column_spec(1, width = "14em") %>%
  column_spec(2, width = "6em") %>%
  column_spec(3, width = "7em") %>%
  footnote(
    general = "MD: mean difference; CrI: credible interval; CHO: carbohydrate; Comp: complex; Prot: Protein",
    # alphabet = c(),
    general_title = "",
    footnote_as_chunk = FALSE
  )

table_n <- tab_inc()
```

<font size = 2> A sensitivity analysis with residual gastric volumes transformed to the log scale to account for skewed distributions yielded consistent results. </font>

<br/>

<font size = 4> Figure `r figure_n`. Forest plot of estimated mean differences in residual gastric volume compared with complex carbohydrate containing liquids. </font>

```{r rgvForest, include = TRUE, fig.align = 'left', fig.width = 5, fig.height = 3}
## (updated 2021/05/15 10:44) forest plot ---------------------------------
nma.forest(re_results, comparator = "CHO/Comp") +
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 12))

figure_n <- fig_inc()
```

<br/>

<hr style="height:2px;border-width:0;color:gray;background-color:gray">

## *Supplemental *


### &emsp;Inconsistency  

<br/>

<font size = 4> Figure `r figure_n`. Posterior mean deviance comparison plot of consistency and inconsistency models (overall no indication of inconsistency).  </font>

```{r rgvConsistencyCalc, include = FALSE, echo = FALSE}
## (updated 2021/07/04 17:20) rgv_inconsistency_re ------------------------
rgv_inconsistency_re <- nma.model(
  data = rgv.data,
  outcome = "mean",
  N = "n",
  sd = "sd",
  reference = "Fasting",
  family = "normal",
  link = "identity",
  effects = "random",
  type = "inconsistency",
  # prior.sigma = "dnorm(0,1) T(0,)"
)

set.seed(12345)
re_inconsist_results <- nma.run(rgv_inconsistency_re,
  n.adapt = 101,
  n.burnin = 20000,
  n.iter = 50000
)

#Plot the results against each other to assess inconsistency
assess.consistency <- nma.fit(re_results)
assess.inconsistency <- nma.fit(re_inconsist_results)
```


```{r rgvConsistency, include = TRUE, fig.align = 'left', fig.width = 7, fig.height = 4}
## (updated 2021/07/04 17:20) rgv_inconsistency_re ------------------------
nma.compare(assess.consistency, assess.inconsistency)
figure_n <- fig_inc()
```

<br/>

<hr style="height:2px;border-width:0;color:gray;background-color:gray">

<br/>

### &emsp;Model Fit 

<br/>

<font size = 4> Figure `r figure_n`. Leverage plot and fit statistics for random-effects model.  </font>

```{r rgvLeverage, include = TRUE, fig.align = 'left', fig.width = 7, fig.height = 4}
## (updated 2021/07/04 17:20) rgv_inconsistency_re ------------------------
assess.consistency <- nma.fit(re_results)
figure_n <- fig_inc()
```

<br/>

<hr style="height:2px;border-width:0;color:gray;background-color:gray">

<br/>

### &emsp;Carbohydrate Amounts  

<br/>

<font size = 4> Figure `r figure_n`. Total carbohydrate consumed prior to surgery according to type and combined with protein for trials in the network meta-analysis. </font>

```{r choAmtsFigure, include = TRUE, echo = FALSE, include = TRUE, fig.width = 6, fig.height = 2.5, fig.align = "left"}
## * (updated 2021/03/10 13:29) cho_amts adult rct ----------------------------
cho_amts <- study_arm_cho.dat %>%
  filter(study %in% rgv_meta.dat$study) %>% 
  filter(design == "rct", age == "Adult", surg_nosurg == "surgical") %>%
  filter(arm_tx %in% c("CHO/Comp/Prot", "CHO/Simp/Prot", "CHO/Comp", "CHO/Simp")) %>%
  filter(refid %in% cho_refids) %>% 
  mutate(
    amtingest2 = replace_na(amtingest2, 0),
    amtingest1 = replace_na(amtingest1, 0),
    tot_vol = amtingest2 + amtingest1,
    prot_gm_all = prot_conc * tot_vol / 100,
    cho_gm_all = cho_conc * tot_vol / 100,
    cho_gm_1 = cho_conc * amtingest1 / 100,
    cho_gm_2 = cho_conc * amtingest2 / 100,
    cho_gm_all = ifelse(is.na(cho_gm_all), cho_gm, cho_gm_all),
    prot_gm_all = ifelse(is.na(prot_gm_all), protein_gm, prot_gm_all)) %>% 
  select(refid, study, arm_tx, arm, cho_gm, protein_gm, prot_gm_all, cho_gm_all, cho_gm_1, cho_gm_2, amtingest1, amtingest2, prot_conc, cho_conc)

## figure of cho amounts ####
theme <- theme_set(theme_minimal())
theme <- theme_update(
  legend.position = "none",
  axis.text = element_text(size = 14), 
  axis.title = element_text(size = 14),
  axis.line.y = element_line(size = 0, colour = "white", linetype = "solid"), 
  axis.ticks = element_line(colour = "gray72"),
  panel.grid.major = element_line(colour = "gray80"))
par(mar = c(6, 5, 2, 2), oma = c(1, 1, 2, 2))

cho_amts %>% ggplot(aes(cho_gm_all, arm_tx)) +
  geom_boxplot(varwidth = TRUE) +
  geom_jitter(width = .2, alpha = 0.8) +
  scale_y_discrete(name = "") +
  scale_x_continuous(name = "Grams", limits = c(0, 200)) 
  # + theme(axis.text=element_text(size = 16),
  #       axis.title=element_text(size = 16,face="bold"))

figure_n <- fig_inc()
```

<font size = 3> CHO: carbohydrate; Comp: complex; Simp: simple; Prot: protein. </font>

<br clear="all" />

<hr style="height:2px;border-width:0;color:gray;background-color:gray">

<br>

# **Risk of Bias**

<br>

<font size = 4> Figure `r figure_n`. Summary risk of bias appraisal for randomized controlled trials included in the analyses.</font>

<img src="assets/cho_nma_rob_summary.svg" style="width:600px;" align="left"/>

<br clear="all" />

```{r}
figure_n <- fig_inc()
```

<br/>

<font size = 4> Figure `r figure_n`. Individual study risk of bias appraisals for randomized controlled trials included in the analyses. </font>

<img src="assets/cho_nma_rob_traffic.svg" style="width:600px;" align="left"/>

```{r}
figure_n <- fig_inc()
```

<br clear="all" />

<hr style="height:2px;border-width:0;color:gray;background-color:gray">

<br/>

# References {#references} 

<br/>
